### 2015-02-19 papers/sbst15/results.tex

com.google.gdata.data.AttributeHelper                     &         1.00    &   9.94    &   0.03    &   92.56      &     92.89      &     62.50 \\
com.google.gdata.data.DateTime                            &         1.00    &   3.62    &   0.02    &   83.31      &     71.67      &     86.25 \\        
com.google.gdata.data.Kind                                &         1.00    &   4.03    &   0.04    &   57.07      &     48.86      &     46.74  \\         
com.google.gdata.data.Link                                &         1.00    &   11.17   &   0.08    &   74.94      &     72.75      &     63.06  \\        
com.google.gdata.data.OtherContent                        &         1.00    &   3.60    &   0.04    &   56.46      &     49.24      &     64.58  \\          
com.google.gdata.data.OutOfLineContent                    &         1.00    &   8.96    &   0.04    &   89.75      &     85.12      &     29.17  \\          
com.google.gdata.data.Source                              &         1.00    &   10.43   &   0.05    &   64.37      &     63.07      &     16.23  \\          
net.sf.javaml.core.AbstractInstance                       &         1.00    &   3.31    &   0.02    &   13.28      &     5.36       &     0.00  \\           
net.sf.javaml.core.Complex                                &         1.00    &   0.28    &   0.02    &   100.00     &     0.00       &     100.00   \\        
net.sf.javaml.core.DefaultDataset                         &         1.00    &   3.87    &   0.02    &   95.31      &     97.50      &     57.80   \\         
net.sf.javaml.core.DenseInstance                          &         1.00    &   3.92    &   0.02    &   98.87      &     100.00     &     100.00   \\        
net.sf.javaml.core.Fold                                   &         1.00    &   4.74    &   0.02    &   91.44      &     96.67      &     79.17   \\         
net.sf.javaml.core.SparseInstance                         &         1.00    &   3.06    &   0.02    &   98.37      &     87.50      &     97.06   \\         
net.sf.javaml.tools.data.ARFFHandler                      &         1.00    &   2.47    &   0.02    &   75.42      &     70.83      &     48.15    \\        
twitter4j.ExceptionDiagnosis                              &         1.00    &   3.81    &   0.01    &   0.00       &     0.00       &     0.00    \\         
twitter4j.GeoQuery                                        &         1.00    &   3.80    &   0.01    &   0.00       &     0.00       &     0.00    \\         
twitter4j.OEmbedRequest                                   &         1.00    &   4.47    &   0.01    &   0.00       &     0.00       &     0.00   \\          
twitter4j.Paging                                          &         1.00    &   3.35    &   0.01    &   0.00       &     0.00       &     0.00   \\          
twitter4j.TwitterBaseImpl                                 &         1.00    &   11.48   &   0.01    &   0.00       &     0.00       &    0.00    \\         
twitter4j.TwitterException                                &         1.00    &   5.12    &   0.01    &   0.00       &     0.00       &     0.00   \\          
twitter4j.TwitterImpl                                     &         1.00    &   16.35   &   0.01    &   0.00       &     0.00       &     0.00    \\         
com.puppycrawl.tools.checkstyle.api.AbstractLoader        &         1.00    &   2.28    &   0.03    &   77.00      &     50.00      &     30.00   \\         
com.puppycrawl.tools.checkstyle.api.AnnotationUtility     &         1.00    &   3.02    &   0.02    &   53.74      &     50.00      &     40.91   \\         
com.puppycrawl.tools.checkstyle.api.AutomaticBean         &         1.00    &   2.75    &   0.04    &   67.50      &     50.00      &     18.60   \\         
com.puppycrawl.tools.checkstyle.api.FileContents          &         1.00    &   5.50    &   0.03    &   97.08      &     91.67      &     80.67    \\        
com.puppycrawl.tools.checkstyle.api.FileText              &         1.00    &   2.94    &   0.03    &   88.62      &     85.26      &     91.84   \\         
com.puppycrawl.tools.checkstyle.api.ScopeUtils            &         1.00    &   14.16   &   0.09    &   72.63      &     50.33      &     24.04   \\         
com.puppycrawl.tools.checkstyle.api.Utils                 &         1.00    &   3.35    &   0.03    &   93.99      &     92.31      &     90.48   \\         
com.google.common.base.CharMatcher                        &         0.00    &   31.32   &   0.00    &   0.00       &     0.00       &     0.00   \\          
com.google.common.base.Joiner                             &         1.00    &   3.37    &   0.03    &   86.52      &     97.10      &     79.79   \\         
com.google.common.base.Objects                            &         1.00    &   2.63    &   0.03    &   98.13      &     95.37      &     94.59    \\        
com.google.common.base.Predicates                         &         1.00    &   4.16    &   0.02    &   44.39      &     21.88      &     30.68   \\         
com.google.common.base.SmallCharMatcher                   &         1.00    &   4.06    &   0.02    &   97.60      &     92.31      &     0.00    \\         
com.google.common.base.Splitter                           &         1.00    &   4.88    &   0.02    &   94.35      &     91.35      &     76.56   \\         
com.google.common.base.Suppliers                          &         1.00    &   3.07    &   0.02    &   60.38      &     58.33      &     56.06   \\         
org.hibernate.search.SearchException                      &         1.00    &   0.26    &   0.01    &   0.00       &     0.00       &     0.00   \\          
org.hibernate.search.Version                              &         1.00    &   0.24    &   0.01    &   0.00       &     0.00       &     0.00   \\          
org.hibernate.search.backend.BackendFactory               &         1.00    &   7.65    &   0.01    &   0.00       &     0.00       &     0.00   \\          
org.hibernate.search.backend.FlushLuceneWork              &         1.00    &   2.28    &   0.01    &   0.00       &     0.00       &     0.00   \\          
org.hibernate.search.backend.OptimizeLuceneWork           &         1.00    &   2.28    &   0.01    &   0.00       &     0.00       &     0.00   \\          
org.hibernate.search.util.logging.impl.LoggerFactory      &         1.00    &   0.27    &   0.01    &   0.00       &     0.00       &     0.00   \\          
org.hibernate.search.util.logging.impl.LoggerHelper       &         1.00    &   0.25    &   0.01    &   0.00       &     0.00       &     0.00   \\          
de.tudarmstadt.ukp.wikipedia.api.CategoryDescendantsIterator &       1.00   &    2.51   &    0.05   &    8.24      &      0.00       &     0.00   \\          
de.tudarmstadt.ukp.wikipedia.api.CycleHandler             &         0.00    &   32.02   &   0.00    &   0.00       &     0.00        &    0.00    \\         
de.tudarmstadt.ukp.wikipedia.api.Page                     &         1.00    &   2.75    &   0.06    &   1.49       &     6.00        &    2.67    \\         
de.tudarmstadt.ukp.wikipedia.api.PageIterator             &         1.00    &   3.76    &   0.06    &   47.72      &     46.83       &    38.54    \\        
de.tudarmstadt.ukp.wikipedia.api.PageQueryIterable        &         1.00    &   3.01    &   0.05    &   18.57      &     9.30        &    0.00    \\         
de.tudarmstadt.ukp.wikipedia.api.Title                    &         1.00    &   2.98    &   0.02    &   89.51      &     83.33       &    98.81   \\         
de.tudarmstadt.ukp.wikipedia.api.WikipediaInfo            &         0.00    &   32.02   &   0.00    &   0.00       &     0.00        &    0.00    \\         
org.asynchttpclient.AsyncHttpClient                       &         1.00    &   6.81    &   0.05    &   67.51      &     63.89       &    63.08   \\         
org.asynchttpclient.AsyncHttpClientConfig                 &         1.00    &   10.24   &   0.06    &   93.73      &     61.11       &    83.03   \\         
org.asynchttpclient.FluentCaseInsensitiveStringsMap       &         1.00    &   5.59    &   0.02    &   98.72      &     95.13       &    87.21   \\         
org.asynchttpclient.FluentStringsMap                      &         1.00    &   4.00    &   0.02    &   98.76      &     94.38       &    86.05   \\         
org.asynchttpclient.Realm                                 &         1.00    &   8.00    &   0.02    &   97.55      &     95.20       &    89.93    \\        
org.asynchttpclient.RequestBuilderBase                    &         1.00    &   8.06    &   0.03    &   93.96      &     89.89       &    43.05   \\         
org.asynchttpclient.SimpleAsyncHttpClient                 &         0.50    &   31.34   &   0.60    &   32.64      &     20.77       &    0.00    \\         
org.scribe.model.OAuthConfig                              &         1.00    &   0.58    &   0.02    &   100.00     &     100.00      &    100.00   \\        
org.scribe.model.OAuthRequest                             &         1.00    &   3.49    &   0.02    &   100.00     &     100.00      &    80.00   \\         
org.scribe.model.ParameterList                            &         1.00    &   2.65    &   0.02    &   99.61      &     99.07       &    91.30    \\        
org.scribe.model.Request                                  &         1.00    &   3.18    &   0.02    &   62.15      &     40.91       &    28.68   \\         
org.scribe.model.Response                                 &         0.00    &   2.20    &   0.00    &   0.00       &     0.00        &    0.00   \\          
org.scribe.model.Token                                    &         1.00    &   2.35    &   0.02    &   100.00     &     100.00      &    77.27   \\         
org.scribe.model.Verifier                                 &         1.00    &   0.12    &   0.02    &   100.00     &     0.00        &    50.00    \\