### 2016-02-21 papers/sbst16/mainTable.tex
added table

\begin{tabular}{ l rrrr rrrr}\toprule 
 Class &  \multicolumn{4}{c}{Branch Coverage}  &  \multicolumn{4}{c}{Fault Detection} \\ 
 & 60s & 120s & 240s & 480s & 60s & 120s & 240s & 480s \\ 
\midrule 
com.google.javascript.jscomp.CheckGlobalThis &  3.4\% &  20.9\% &  20.5\% &  26.5\% &  0.0\% &  0.0\% &  0.0\% &  8.3\%\\ 
com.google.javascript.jscomp.CollapseProperties &  2.6\% &  3.6\% &  5.3\% &  6.5\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
com.google.javascript.jscomp.ControlFlowAnalysis &  1.1\% &  1.4\% &  20.4\% &  25.9\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
com.google.javascript.jscomp.ExploitAssigns &  0.0\% &  2.5\% &  22.3\% &  25.7\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
com.google.javascript.jscomp.PeepholeFoldConstants &  0.2\% &  0.2\% &  5.0\% &  6.5\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
com.google.javascript.jscomp.PeepholeSubstituteAlternateSyntax &  3.6\% &  4.7\% &  6.8\% &  9.6\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
com.google.javascript.jscomp.ReferenceCollectingCallback &  5.8\% &  9.0\% &  20.6\% &  26.3\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
com.google.javascript.jscomp.ScopedAliases &  0.0\% &  5.6\% &  11.3\% &  11.8\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
com.google.javascript.jscomp.parsing.JsDocInfoParser &  11.3\% &  12.4\% &  16.5\% &  24.6\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
com.google.javascript.rhino.jstype.RecordType &  0.0\% &  0.0\% &  63.7\% &  75.7\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
org.apache.commons.lang.ClassUtils &  61.3\% &  65.6\% &  67.7\% &  59.2\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
org.apache.commons.lang.LocaleUtils &  75.0\% &  70.8\% &  66.4\% &  70.1\% &  83.3\% &  100.0\% &  100.0\% &  100.0\%\\ 
org.apache.commons.lang.math.NumberUtils &  64.1\% &  65.9\% &  73.4\% &  77.7\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.apache.commons.lang.text.ExtendedMessageFormat &  32.0\% &  37.9\% &  41.6\% &  54.4\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
org.apache.commons.lang.text.StrBuilder &  66.6\% &  72.2\% &  78.3\% &  82.6\% &  33.3\% &  38.9\% &  55.6\% &  61.1\%\\ 
org.apache.commons.lang.time.DateUtils &  56.6\% &  81.3\% &  87.9\% &  93.6\% &  33.3\% &  16.7\% &  83.3\% &  100.0\%\\ 
org.apache.commons.lang.time.DurationFormatUtils &  78.8\% &  82.6\% &  80.2\% &  90.1\% &  0.0\% &  0.0\% &  0.0\% &  16.7\%\\ 
org.apache.commons.lang.time.FastDateFormat &  63.2\% &  43.4\% &  73.8\% &  73.0\% &  0.0\% &  0.0\% &  0.0\% &  16.7\%\\ 
org.apache.commons.lang3.ArrayUtils &  61.5\% &  66.5\% &  72.2\% &  75.1\% &  16.7\% &  0.0\% &  0.0\% &  33.3\%\\ 
org.apache.commons.lang3.ClassUtils &  60.9\% &  61.0\% &  66.7\% &  59.9\% &  100.0\% &  100.0\% &  100.0\% &  83.3\%\\ 
org.apache.commons.lang3.math.NumberUtils &  67.1\% &  70.8\% &  76.0\% &  80.0\% &  16.7\% &  0.0\% &  83.3\% &  83.3\%\\ 
org.apache.commons.lang3.text.translate.NumericEntityUnescaper &  6.2\% &  27.1\% &  51.0\% &  79.2\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.apache.commons.math.distribution.NormalDistributionImpl &  77.8\% &  77.8\% &  77.8\% &  77.8\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.apache.commons.math.fraction.Fraction &  94.1\% &  95.9\% &  97.4\% &  98.5\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.apache.commons.math.fraction.ProperFractionFormat &  46.5\% &  70.2\% &  76.3\% &  70.2\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
org.apache.commons.math.geometry.euclidean.threed.Rotation &  45.5\% &  59.0\% &  88.7\% &  91.5\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.apache.commons.math.ode.AbstractIntegrator &  49.6\% &  50.7\% &  58.3\% &  64.5\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.apache.commons.math.ode.nonstiff.EmbeddedRungeKuttaIntegrator &  39.2\% &  36.1\% &  69.1\% &  71.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.apache.commons.math.optimization.MultiStartUnivariateRealOptimizer &  81.9\% &  90.7\% &  86.8\% &  97.1\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
org.apache.commons.math.optimization.general.LevenbergMarquardtOptimizer &  22.7\% &  27.6\% &  55.0\% &  64.8\% &  0.0\% &  66.7\% &  66.7\% &  83.3\%\\ 
org.apache.commons.math.optimization.linear.SimplexTableau &  2.2\% &  74.8\% &  86.9\% &  96.1\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
org.apache.commons.math.util.MathUtils &  78.9\% &  80.4\% &  83.5\% &  85.9\% &  0.0\% &  16.7\% &  0.0\% &  16.7\%\\ 
org.apache.commons.math.util.MultidimensionalCounter &  94.3\% &  95.8\% &  98.4\% &  98.4\% &  16.7\% &  16.7\% &  83.3\% &  100.0\%\\ 
org.apache.commons.math3.distribution.HypergeometricDistribution &  96.8\% &  96.2\% &  98.7\% &  99.4\% &  0.0\% &  16.7\% &  0.0\% &  0.0\%\\ 
org.apache.commons.math3.linear.RectangularCholeskyDecomposition &  58.3\% &  70.2\% &  73.2\% &  91.1\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.apache.commons.math3.ode.AbstractIntegrator &  29.6\% &  39.4\% &  59.2\% &  63.8\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.apache.commons.math3.optimization.direct.CMAESOptimizer &  69.3\% &  72.0\% &  69.3\% &  71.3\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.jfree.chart.axis.Axis &  37.5\% &  56.9\% &  63.2\% &  57.1\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
org.jfree.chart.plot.MultiplePiePlot &  48.0\% &  53.4\% &  55.6\% &  45.8\% &  16.7\% &  16.7\% &  0.0\% &  0.0\%\\ 
org.jfree.chart.plot.ValueMarker &  87.5\% &  87.5\% &  87.5\% &  75.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.jfree.chart.plot.XYPlot &  29.5\% &  32.5\% &  41.6\% &  47.6\% &  16.7\% &  0.0\% &  100.0\% &  100.0\%\\ 
org.jfree.chart.renderer.GrayPaintScale &  100.0\% &  100.0\% &  100.0\% &  100.0\% &  100.0\% &  33.3\% &  50.0\% &  66.7\%\\ 
org.jfree.chart.renderer.category.AbstractCategoryItemRenderer &  22.6\% &  26.4\% &  39.9\% &  45.7\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
org.jfree.chart.renderer.category.MinMaxCategoryRenderer &  30.7\% &  36.6\% &  34.8\% &  26.2\% &  0.0\% &  83.3\% &  50.0\% &  66.7\%\\ 
org.jfree.chart.renderer.category.StatisticalBarRenderer &  8.3\% &  8.9\% &  10.0\% &  16.5\% &  0.0\% &  0.0\% &  50.0\% &  50.0\%\\ 
org.jfree.chart.util.ShapeList &  37.5\% &  49.0\% &  49.0\% &  50.0\% &  83.3\% &  100.0\% &  83.3\% &  100.0\%\\ 
org.jfree.chart.util.ShapeUtilities &  58.0\% &  82.8\% &  57.3\% &  76.6\% &  50.0\% &  100.0\% &  66.7\% &  83.3\%\\ 
org.jfree.data.category.DefaultIntervalCategoryDataset &  74.6\% &  78.1\% &  80.1\% &  85.0\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
org.jfree.data.general.DatasetUtilities &  0.0\% &  15.9\% &  62.2\% &  66.9\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.jfree.data.time.TimePeriodValues &  38.3\% &  89.2\% &  95.4\% &  98.5\% &  0.0\% &  16.7\% &  50.0\% &  66.7\%\\ 
org.jfree.data.time.TimeSeries &  57.6\% &  69.6\% &  79.3\% &  86.4\% &  33.3\% &  44.4\% &  44.4\% &  50.0\%\\ 
org.joda.time.DateTimeZone &  69.9\% &  71.6\% &  73.4\% &  77.6\% &  33.3\% &  50.0\% &  33.3\% &  50.0\%\\ 
org.joda.time.MutableDateTime &  79.6\% &  84.4\% &  90.5\% &  96.8\% &  100.0\% &  50.0\% &  33.3\% &  16.7\%\\ 
org.joda.time.Partial &  55.5\% &  61.0\% &  87.1\% &  93.6\% &  16.7\% &  0.0\% &  50.0\% &  33.3\%\\ 
org.joda.time.Period &  81.0\% &  84.4\% &  90.6\% &  93.2\% &  33.3\% &  50.0\% &  83.3\% &  100.0\%\\ 
org.joda.time.base.BaseSingleFieldPeriod &  87.3\% &  90.6\% &  86.6\% &  92.4\% &  0.0\% &  0.0\% &  16.7\% &  16.7\%\\ 
org.joda.time.format.DateTimeFormatter &  84.1\% &  80.3\% &  84.1\% &  87.5\% &  33.3\% &  16.7\% &  16.7\% &  0.0\%\\ 
org.joda.time.format.DateTimeFormatterBuilder &  25.3\% &  25.5\% &  52.6\% &  60.9\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
org.joda.time.format.PeriodFormatterBuilder &  29.0\% &  31.0\% &  44.6\% &  54.5\% &  0.0\% &  0.0\% &  0.0\% &  16.7\%\\ 
org.joda.time.tz.ZoneInfoCompiler &  25.9\% &  27.9\% &  40.2\% &  30.7\% &  100.0\% &  100.0\% &  100.0\% &  66.7\%\\ 
\bottomrule 
\end{tabular}