### 2016-02-21 papers/sbst16/coverageTable.tex
Update tables and paper

\begin{tabular}{ ll rrrr rrrr}\toprule 
 Defects4J ID & Class &  \multicolumn{4}{c}{Line Coverage}  &  \multicolumn{4}{c}{Branch Coverage} \\ 
 & & 60s & 120s & 240s & 480s & 60s & 120s & 240s & 480s \\ 
\midrule 
Chart-1 &  org.jfree.chart.renderer.category.AbstractCategoryItemRenderer &  34.5\% &  36.8\% &  51.0\% &  56.2\% &  22.6\% &  26.4\% &  39.9\% &  45.7\%\\ 
Chart-11 &  org.jfree.chart.util.ShapeUtilities &  81.0\% &  92.1\% &  62.1\% &  79.4\% &  58.0\% &  82.8\% &  57.3\% &  76.6\%\\ 
Chart-12 &  org.jfree.chart.plot.MultiplePiePlot &  54.2\% &  62.0\% &  64.9\% &  51.3\% &  48.0\% &  53.4\% &  55.6\% &  45.8\%\\ 
Chart-16 &  org.jfree.data.category.DefaultIntervalCategoryDataset &  81.8\% &  82.7\% &  85.2\% &  88.2\% &  74.6\% &  78.1\% &  80.1\% &  85.0\%\\ 
Chart-17 &  org.jfree.data.time.TimeSeries &  69.2\% &  83.0\% &  87.3\% &  89.5\% &  58.9\% &  74.3\% &  80.9\% &  84.5\%\\ 
Chart-2 &  org.jfree.data.general.DatasetUtilities &  0.0\% &  18.0\% &  67.2\% &  71.4\% &  0.0\% &  15.9\% &  62.2\% &  66.9\%\\ 
Chart-20 &  org.jfree.chart.plot.ValueMarker &  95.5\% &  95.5\% &  95.5\% &  80.3\% &  87.5\% &  87.5\% &  87.5\% &  75.0\%\\ 
Chart-23 &  org.jfree.chart.renderer.category.MinMaxCategoryRenderer &  41.6\% &  44.2\% &  43.1\% &  30.8\% &  30.7\% &  36.6\% &  34.8\% &  26.2\%\\ 
Chart-24 &  org.jfree.chart.renderer.GrayPaintScale &  100.0\% &  100.0\% &  100.0\% &  100.0\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
Chart-25 &  org.jfree.chart.renderer.category.StatisticalBarRenderer &  13.1\% &  13.4\% &  18.4\% &  26.8\% &  8.3\% &  8.9\% &  10.0\% &  16.5\%\\ 
Chart-26 &  org.jfree.chart.axis.Axis &  54.6\% &  63.7\% &  73.5\% &  65.7\% &  37.5\% &  56.9\% &  63.2\% &  57.1\%\\ 
Chart-3 &  org.jfree.data.time.TimeSeries &  77.5\% &  78.8\% &  90.3\% &  93.2\% &  60.5\% &  64.7\% &  81.4\% &  87.5\%\\ 
Chart-4 &  org.jfree.chart.plot.XYPlot &  45.5\% &  48.1\% &  55.9\% &  60.8\% &  29.5\% &  32.5\% &  41.6\% &  47.6\%\\ 
Chart-6 &  org.jfree.chart.util.ShapeList &  51.0\% &  54.9\% &  55.4\% &  55.4\% &  37.5\% &  49.0\% &  49.0\% &  50.0\%\\ 
Chart-7 &  org.jfree.data.time.TimePeriodValues &  58.6\% &  95.3\% &  97.0\% &  98.1\% &  38.3\% &  89.2\% &  95.4\% &  98.5\%\\ 
Chart-9 &  org.jfree.data.time.TimeSeries &  65.8\% &  79.8\% &  83.6\% &  92.3\% &  53.3\% &  69.8\% &  75.6\% &  87.0\%\\ 
Closure-100 &  com.google.javascript.jscomp.CheckGlobalThis &  24.4\% &  42.6\% &  40.7\% &  51.1\% &  3.6\% &  21.7\% &  21.1\% &  29.2\%\\ 
Closure-124 &  com.google.javascript.jscomp.ExploitAssigns &  9.5\% &  12.8\% &  36.7\% &  39.9\% &  0.0\% &  2.5\% &  22.3\% &  25.7\%\\ 
Closure-130 &  com.google.javascript.jscomp.CollapseProperties &  9.1\% &  11.3\% &  13.1\% &  15.4\% &  2.6\% &  3.6\% &  5.3\% &  6.5\%\\ 
Closure-132 &  com.google.javascript.jscomp.PeepholeSubstituteAlternateSyntax &  9.8\% &  13.4\% &  11.3\% &  14.1\% &  3.1\% &  4.8\% &  6.3\% &  8.4\%\\ 
Closure-14 &  com.google.javascript.jscomp.ControlFlowAnalysis &  9.4\% &  10.5\% &  31.9\% &  37.6\% &  1.1\% &  1.4\% &  20.4\% &  25.9\%\\ 
Closure-16 &  com.google.javascript.jscomp.ScopedAliases &  12.4\% &  18.3\% &  24.2\% &  24.2\% &  0.0\% &  5.6\% &  11.3\% &  11.8\%\\ 
Closure-20 &  com.google.javascript.jscomp.PeepholeSubstituteAlternateSyntax &  11.0\% &  12.8\% &  12.7\% &  17.9\% &  4.0\% &  4.7\% &  7.2\% &  10.8\%\\ 
Closure-46 &  com.google.javascript.rhino.jstype.RecordType &  2.3\% &  2.3\% &  75.6\% &  91.1\% &  0.0\% &  0.0\% &  63.7\% &  75.7\%\\ 
Closure-68 &  com.google.javascript.jscomp.parsing.JsDocInfoParser &  16.6\% &  17.4\% &  22.9\% &  33.5\% &  11.3\% &  12.4\% &  16.5\% &  24.6\%\\ 
Closure-74 &  com.google.javascript.jscomp.PeepholeFoldConstants &  5.6\% &  5.8\% &  8.9\% &  10.9\% &  0.2\% &  0.2\% &  5.0\% &  6.5\%\\ 
Closure-98 &  com.google.javascript.jscomp.ReferenceCollectingCallback &  29.2\% &  35.3\% &  57.3\% &  63.0\% &  5.8\% &  9.0\% &  20.6\% &  26.3\%\\ 
Closure-99 &  com.google.javascript.jscomp.CheckGlobalThis &  24.4\% &  43.3\% &  42.6\% &  48.5\% &  3.2\% &  20.2\% &  19.9\% &  23.9\%\\ 
Lang-28 &  org.apache.commons.lang3.text.translate.NumericEntityUnescaper &  12.0\% &  23.3\% &  47.3\% &  75.3\% &  6.2\% &  27.1\% &  51.0\% &  79.2\%\\ 
Lang-33 &  org.apache.commons.lang3.ClassUtils &  80.7\% &  81.9\% &  84.6\% &  77.0\% &  60.9\% &  61.0\% &  66.7\% &  59.9\%\\ 
Lang-36 &  org.apache.commons.lang3.math.NumberUtils &  83.9\% &  86.2\% &  90.3\% &  92.1\% &  67.1\% &  70.8\% &  76.0\% &  80.0\%\\ 
Lang-37 &  org.apache.commons.lang3.ArrayUtils &  78.6\% &  83.5\% &  88.1\% &  90.1\% &  61.5\% &  66.5\% &  72.2\% &  75.1\%\\ 
Lang-41 &  org.apache.commons.lang.ClassUtils &  81.4\% &  84.8\% &  86.1\% &  76.2\% &  61.3\% &  65.6\% &  67.7\% &  59.2\%\\ 
Lang-43 &  org.apache.commons.lang.text.ExtendedMessageFormat &  49.6\% &  54.1\% &  53.3\% &  62.9\% &  32.0\% &  37.9\% &  41.6\% &  54.4\%\\ 
Lang-47 &  org.apache.commons.lang.text.StrBuilder &  75.7\% &  82.0\% &  86.2\% &  90.6\% &  63.8\% &  72.3\% &  78.7\% &  84.6\%\\ 
Lang-50 &  org.apache.commons.lang.time.FastDateFormat &  80.4\% &  53.6\% &  85.1\% &  86.3\% &  63.2\% &  43.4\% &  73.8\% &  73.0\%\\ 
Lang-57 &  org.apache.commons.lang.LocaleUtils &  86.2\% &  86.4\% &  84.0\% &  85.7\% &  75.0\% &  70.8\% &  66.4\% &  70.1\%\\ 
Lang-58 &  org.apache.commons.lang.math.NumberUtils &  79.4\% &  82.3\% &  86.3\% &  89.8\% &  64.1\% &  65.9\% &  73.4\% &  77.7\%\\ 
Lang-59 &  org.apache.commons.lang.text.StrBuilder &  78.5\% &  79.9\% &  86.0\% &  89.3\% &  68.8\% &  71.4\% &  79.1\% &  82.0\%\\ 
Lang-60 &  org.apache.commons.lang.text.StrBuilder &  77.3\% &  80.9\% &  85.5\% &  88.6\% &  67.2\% &  72.9\% &  77.0\% &  81.3\%\\ 
Lang-63 &  org.apache.commons.lang.time.DurationFormatUtils &  81.1\% &  84.8\% &  82.6\% &  92.1\% &  78.8\% &  82.6\% &  80.2\% &  90.1\%\\ 
Lang-65 &  org.apache.commons.lang.time.DateUtils &  72.5\% &  89.3\% &  94.1\% &  97.5\% &  56.6\% &  81.3\% &  87.9\% &  93.6\%\\ 
Math-103 &  org.apache.commons.math.distribution.NormalDistributionImpl &  85.7\% &  85.7\% &  85.7\% &  85.7\% &  77.8\% &  77.8\% &  77.8\% &  77.8\%\\ 
Math-106 &  org.apache.commons.math.fraction.ProperFractionFormat &  63.5\% &  80.2\% &  85.4\% &  79.9\% &  46.5\% &  70.2\% &  76.3\% &  70.2\%\\ 
Math-18 &  org.apache.commons.math3.optimization.direct.CMAESOptimizer &  85.7\% &  86.1\% &  82.3\% &  84.2\% &  71.0\% &  72.2\% &  66.7\% &  69.6\%\\ 
Math-2 &  org.apache.commons.math3.distribution.HypergeometricDistribution &  98.5\% &  98.2\% &  99.7\% &  99.5\% &  96.8\% &  96.2\% &  98.7\% &  99.4\%\\ 
Math-20 &  org.apache.commons.math3.optimization.direct.CMAESOptimizer &  85.4\% &  87.3\% &  86.7\% &  87.3\% &  67.7\% &  71.8\% &  71.9\% &  72.9\%\\ 
Math-21 &  org.apache.commons.math3.linear.RectangularCholeskyDecomposition &  58.5\% &  69.7\% &  73.3\% &  91.5\% &  58.3\% &  70.2\% &  73.2\% &  91.1\%\\ 
Math-39 &  org.apache.commons.math.ode.nonstiff.EmbeddedRungeKuttaIntegrator &  74.4\% &  67.1\% &  89.5\% &  90.9\% &  39.2\% &  36.1\% &  69.1\% &  71.0\%\\ 
Math-44 &  org.apache.commons.math.ode.AbstractIntegrator &  63.1\% &  62.6\% &  68.2\% &  73.3\% &  49.6\% &  50.7\% &  58.3\% &  64.5\%\\ 
Math-52 &  org.apache.commons.math.geometry.euclidean.threed.Rotation &  72.9\% &  81.3\% &  97.0\% &  98.7\% &  45.5\% &  59.0\% &  88.7\% &  91.5\%\\ 
Math-56 &  org.apache.commons.math.util.MultidimensionalCounter &  97.9\% &  99.3\% &  99.8\% &  99.8\% &  94.3\% &  95.8\% &  98.4\% &  98.4\%\\ 
Math-64 &  org.apache.commons.math.optimization.general.LevenbergMarquardtOptimizer &  35.8\% &  40.9\% &  66.5\% &  76.9\% &  22.7\% &  27.6\% &  55.0\% &  64.8\%\\ 
Math-67 &  org.apache.commons.math.optimization.MultiStartUnivariateRealOptimizer &  87.6\% &  88.7\% &  86.9\% &  90.0\% &  81.9\% &  90.7\% &  86.8\% &  97.1\%\\ 
Math-7 &  org.apache.commons.math3.ode.AbstractIntegrator &  50.0\% &  58.9\% &  74.7\% &  77.6\% &  29.6\% &  39.4\% &  59.2\% &  63.8\%\\ 
Math-88 &  org.apache.commons.math.optimization.linear.SimplexTableau &  8.4\% &  84.0\% &  94.3\% &  96.1\% &  2.2\% &  74.8\% &  86.9\% &  96.1\%\\ 
Math-91 &  org.apache.commons.math.fraction.Fraction &  96.5\% &  97.2\% &  97.5\% &  97.9\% &  94.1\% &  95.9\% &  97.4\% &  98.5\%\\ 
Math-93 &  org.apache.commons.math.util.MathUtils &  81.6\% &  84.1\% &  86.0\% &  87.5\% &  78.9\% &  80.4\% &  83.5\% &  85.9\%\\ 
Time-10 &  org.joda.time.base.BaseSingleFieldPeriod &  91.4\% &  91.4\% &  88.6\% &  91.2\% &  87.3\% &  90.6\% &  86.6\% &  92.4\%\\ 
Time-11 &  org.joda.time.tz.ZoneInfoCompiler &  30.7\% &  32.5\% &  46.3\% &  34.0\% &  25.9\% &  27.9\% &  40.2\% &  30.7\%\\ 
Time-13 &  org.joda.time.format.PeriodFormatterBuilder &  49.8\% &  52.7\% &  65.5\% &  72.7\% &  29.0\% &  31.0\% &  44.6\% &  54.5\%\\ 
Time-20 &  org.joda.time.format.DateTimeFormatterBuilder &  35.3\% &  36.5\% &  68.5\% &  77.0\% &  25.3\% &  25.5\% &  52.6\% &  60.9\%\\ 
Time-23 &  org.joda.time.DateTimeZone &  83.6\% &  83.9\% &  84.0\% &  85.4\% &  71.5\% &  72.7\% &  73.9\% &  77.5\%\\ 
Time-3 &  org.joda.time.MutableDateTime &  87.4\% &  91.0\% &  93.6\% &  95.4\% &  79.6\% &  84.4\% &  90.5\% &  96.8\%\\ 
Time-4 &  org.joda.time.Partial &  72.5\% &  77.0\% &  93.5\% &  97.8\% &  55.5\% &  61.0\% &  87.1\% &  93.6\%\\ 
Time-5 &  org.joda.time.Period &  96.7\% &  97.3\% &  98.4\% &  98.9\% &  81.0\% &  84.4\% &  90.6\% &  93.2\%\\ 
Time-7 &  org.joda.time.format.DateTimeFormatter &  87.3\% &  86.8\% &  90.1\% &  91.7\% &  84.1\% &  80.3\% &  84.1\% &  87.5\%\\ 
Time-8 &  org.joda.time.DateTimeZone &  81.3\% &  81.4\% &  82.8\% &  84.8\% &  68.3\% &  70.4\% &  72.9\% &  77.7\%\\ 
\midrule 
Average &  &  58.3\% &  63.7\% &  71.1\% &  73.9\% &  46.6\% &  53.5\% &  61.7\% &  65.6\%\\ 
\bottomrule 
\end{tabular}