### 2016-02-21 papers/sbst16/faultTable.tex

\begin{tabular}{ ll rrrr rrrr}\toprule 
 Defects4J ID & Class &  \multicolumn{4}{c}{Mutation Score}  &  \multicolumn{4}{c}{Fault Detection} \\ 
 & & 60s & 120s & 240s & 480s & 60s & 120s & 240s & 480s \\ 
\midrule 
Chart-1 &  org.jfree.chart.renderer.category.AbstractCategoryItemRenderer &  11.3\% &  12.2\% &  17.4\% &  22.4\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
Chart-11 &  org.jfree.chart.util.ShapeUtilities &  13.4\% &  28.0\% &  18.9\% &  25.4\% &  50.0\% &  100.0\% &  66.7\% &  83.3\%\\ 
Chart-12 &  org.jfree.chart.plot.MultiplePiePlot &  10.8\% &  19.6\% &  20.3\% &  17.7\% &  16.7\% &  16.7\% &  0.0\% &  0.0\%\\ 
Chart-16 &  org.jfree.data.category.DefaultIntervalCategoryDataset &  47.3\% &  48.6\% &  52.7\% &  55.4\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
Chart-17 &  org.jfree.data.time.TimeSeries &  27.9\% &  32.7\% &  39.0\% &  42.4\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
Chart-2 &  org.jfree.data.general.DatasetUtilities &  0.0\% &  7.3\% &  43.4\% &  49.9\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Chart-20 &  org.jfree.chart.plot.ValueMarker &  81.8\% &  81.8\% &  81.8\% &  69.7\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Chart-23 &  org.jfree.chart.renderer.category.MinMaxCategoryRenderer &  3.4\% &  8.4\% &  6.9\% &  6.3\% &  0.0\% &  83.3\% &  50.0\% &  66.7\%\\ 
Chart-24 &  org.jfree.chart.renderer.GrayPaintScale &  91.9\% &  65.3\% &  73.4\% &  82.9\% &  100.0\% &  33.3\% &  50.0\% &  66.7\%\\ 
Chart-25 &  org.jfree.chart.renderer.category.StatisticalBarRenderer &  0.3\% &  1.9\% &  2.4\% &  2.3\% &  0.0\% &  0.0\% &  50.0\% &  50.0\%\\ 
Chart-26 &  org.jfree.chart.axis.Axis &  5.7\% &  17.1\% &  19.8\% &  17.5\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
Chart-3 &  org.jfree.data.time.TimeSeries &  28.0\% &  33.8\% &  44.9\% &  47.8\% &  0.0\% &  16.7\% &  33.3\% &  50.0\%\\ 
Chart-4 &  org.jfree.chart.plot.XYPlot &  17.3\% &  17.2\% &  19.8\% &  36.6\% &  16.7\% &  0.0\% &  100.0\% &  100.0\%\\ 
Chart-6 &  org.jfree.chart.util.ShapeList &  15.5\% &  23.8\% &  19.0\% &  20.2\% &  83.3\% &  100.0\% &  83.3\% &  100.0\%\\ 
Chart-7 &  org.jfree.data.time.TimePeriodValues &  21.7\% &  63.9\% &  67.3\% &  61.2\% &  0.0\% &  16.7\% &  50.0\% &  66.7\%\\ 
Chart-9 &  org.jfree.data.time.TimeSeries &  23.0\% &  31.5\% &  36.2\% &  44.1\% &  0.0\% &  16.7\% &  0.0\% &  0.0\%\\ 
Closure-100 &  com.google.javascript.jscomp.CheckGlobalThis &  8.2\% &  18.2\% &  20.6\% &  23.6\% &  0.0\% &  0.0\% &  0.0\% &  16.7\%\\ 
Closure-124 &  com.google.javascript.jscomp.ExploitAssigns &  1.4\% &  2.5\% &  15.1\% &  16.9\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Closure-130 &  com.google.javascript.jscomp.CollapseProperties &  0.0\% &  0.4\% &  0.7\% &  1.3\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Closure-132 &  com.google.javascript.jscomp.PeepholeSubstituteAlternateSyntax &  3.5\% &  8.8\% &  7.4\% &  13.7\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Closure-14 &  com.google.javascript.jscomp.ControlFlowAnalysis &  2.7\% &  3.6\% &  20.0\% &  26.3\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Closure-16 &  com.google.javascript.jscomp.ScopedAliases &  0.0\% &  4.0\% &  5.6\% &  4.4\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Closure-20 &  com.google.javascript.jscomp.PeepholeSubstituteAlternateSyntax &  3.9\% &  8.5\% &  10.0\% &  13.4\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Closure-46 &  com.google.javascript.rhino.jstype.RecordType &  0.0\% &  0.0\% &  37.4\% &  43.1\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
Closure-68 &  com.google.javascript.jscomp.parsing.JsDocInfoParser &  3.7\% &  4.4\% &  9.3\% &  13.5\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Closure-74 &  com.google.javascript.jscomp.PeepholeFoldConstants &  1.9\% &  2.0\% &  2.5\% &  3.2\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Closure-98 &  com.google.javascript.jscomp.ReferenceCollectingCallback &  8.3\% &  9.1\% &  22.4\% &  27.6\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Closure-99 &  com.google.javascript.jscomp.CheckGlobalThis &  7.0\% &  17.0\% &  18.2\% &  21.1\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Lang-28 &  org.apache.commons.lang3.text.translate.NumericEntityUnescaper &  8.3\% &  17.7\% &  24.2\% &  49.8\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Lang-33 &  org.apache.commons.lang3.ClassUtils &  36.2\% &  39.4\% &  43.7\% &  38.1\% &  100.0\% &  100.0\% &  100.0\% &  83.3\%\\ 
Lang-36 &  org.apache.commons.lang3.math.NumberUtils &  27.1\% &  36.2\% &  38.7\% &  41.8\% &  16.7\% &  0.0\% &  83.3\% &  83.3\%\\ 
Lang-37 &  org.apache.commons.lang3.ArrayUtils &  24.4\% &  35.0\% &  56.6\% &  58.2\% &  16.7\% &  0.0\% &  0.0\% &  33.3\%\\ 
Lang-41 &  org.apache.commons.lang.ClassUtils &  39.1\% &  44.2\% &  46.5\% &  36.8\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
Lang-43 &  org.apache.commons.lang.text.ExtendedMessageFormat &  12.4\% &  16.6\% &  19.4\% &  25.8\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
Lang-47 &  org.apache.commons.lang.text.StrBuilder &  23.2\% &  32.2\% &  44.0\% &  54.6\% &  66.7\% &  66.7\% &  66.7\% &  83.3\%\\ 
Lang-50 &  org.apache.commons.lang.time.FastDateFormat &  19.2\% &  22.9\% &  50.8\% &  56.5\% &  0.0\% &  0.0\% &  0.0\% &  16.7\%\\ 
Lang-57 &  org.apache.commons.lang.LocaleUtils &  40.5\% &  39.2\% &  37.7\% &  38.4\% &  83.3\% &  100.0\% &  100.0\% &  100.0\%\\ 
Lang-58 &  org.apache.commons.lang.math.NumberUtils &  21.1\% &  33.3\% &  37.6\% &  42.5\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Lang-59 &  org.apache.commons.lang.text.StrBuilder &  25.3\% &  31.1\% &  44.8\% &  49.9\% &  16.7\% &  33.3\% &  83.3\% &  66.7\%\\ 
Lang-60 &  org.apache.commons.lang.text.StrBuilder &  23.8\% &  34.0\% &  44.9\% &  48.6\% &  16.7\% &  16.7\% &  16.7\% &  33.3\%\\ 
Lang-63 &  org.apache.commons.lang.time.DurationFormatUtils &  13.6\% &  20.4\% &  20.8\% &  25.5\% &  0.0\% &  0.0\% &  0.0\% &  16.7\%\\ 
Lang-65 &  org.apache.commons.lang.time.DateUtils &  28.1\% &  35.7\% &  51.9\% &  55.3\% &  33.3\% &  16.7\% &  83.3\% &  100.0\%\\ 
Math-103 &  org.apache.commons.math.distribution.NormalDistributionImpl &  63.0\% &  62.8\% &  63.7\% &  63.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Math-106 &  org.apache.commons.math.fraction.ProperFractionFormat &  25.0\% &  32.2\% &  40.9\% &  35.1\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
Math-18 &  org.apache.commons.math3.optimization.direct.CMAESOptimizer &  29.9\% &  29.0\% &  26.5\% &  26.5\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Math-2 &  org.apache.commons.math3.distribution.HypergeometricDistribution &  39.3\% &  84.8\% &  89.5\% &  90.8\% &  0.0\% &  16.7\% &  0.0\% &  0.0\%\\ 
Math-20 &  org.apache.commons.math3.optimization.direct.CMAESOptimizer &  33.1\% &  35.9\% &  28.7\% &  27.3\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Math-21 &  org.apache.commons.math3.linear.RectangularCholeskyDecomposition &  23.7\% &  32.5\% &  28.1\% &  44.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Math-39 &  org.apache.commons.math.ode.nonstiff.EmbeddedRungeKuttaIntegrator &  20.7\% &  17.2\% &  30.3\% &  35.6\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Math-44 &  org.apache.commons.math.ode.AbstractIntegrator &  35.0\% &  32.6\% &  35.0\% &  41.4\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Math-52 &  org.apache.commons.math.geometry.euclidean.threed.Rotation &  52.8\% &  58.8\% &  79.5\% &  82.8\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Math-56 &  org.apache.commons.math.util.MultidimensionalCounter &  48.8\% &  56.2\% &  74.3\% &  74.4\% &  16.7\% &  16.7\% &  83.3\% &  100.0\%\\ 
Math-64 &  org.apache.commons.math.optimization.general.LevenbergMarquardtOptimizer &  5.7\% &  5.9\% &  16.4\% &  24.0\% &  0.0\% &  66.7\% &  66.7\% &  83.3\%\\ 
Math-67 &  org.apache.commons.math.optimization.MultiStartUnivariateRealOptimizer &  36.2\% &  52.2\% &  47.4\% &  59.4\% &  100.0\% &  100.0\% &  100.0\% &  100.0\%\\ 
Math-7 &  org.apache.commons.math3.ode.AbstractIntegrator &  26.2\% &  32.5\% &  33.1\% &  31.5\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Math-88 &  org.apache.commons.math.optimization.linear.SimplexTableau &  0.5\% &  41.6\% &  52.8\% &  57.1\% &  0.0\% &  0.0\% &  16.7\% &  0.0\%\\ 
Math-91 &  org.apache.commons.math.fraction.Fraction &  58.7\% &  69.1\% &  70.8\% &  70.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Math-93 &  org.apache.commons.math.util.MathUtils &  24.6\% &  46.0\% &  65.7\% &  68.8\% &  0.0\% &  16.7\% &  0.0\% &  16.7\%\\ 
Time-10 &  org.joda.time.base.BaseSingleFieldPeriod &  46.3\% &  63.3\% &  59.6\% &  63.4\% &  0.0\% &  0.0\% &  16.7\% &  16.7\%\\ 
Time-11 &  org.joda.time.tz.ZoneInfoCompiler &  11.4\% &  15.7\% &  19.4\% &  14.6\% &  100.0\% &  100.0\% &  100.0\% &  66.7\%\\ 
Time-13 &  org.joda.time.format.PeriodFormatterBuilder &  11.9\% &  16.8\% &  25.1\% &  30.9\% &  0.0\% &  0.0\% &  0.0\% &  16.7\%\\ 
Time-20 &  org.joda.time.format.DateTimeFormatterBuilder &  9.3\% &  11.2\% &  27.7\% &  39.4\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
Time-23 &  org.joda.time.DateTimeZone &  40.4\% &  49.3\% &  50.2\% &  51.7\% &  0.0\% &  16.7\% &  0.0\% &  0.0\%\\ 
Time-3 &  org.joda.time.MutableDateTime &  35.4\% &  47.6\% &  55.9\% &  57.6\% &  100.0\% &  50.0\% &  33.3\% &  16.7\%\\ 
Time-4 &  org.joda.time.Partial &  29.6\% &  38.7\% &  67.9\% &  73.9\% &  16.7\% &  0.0\% &  50.0\% &  33.3\%\\ 
Time-5 &  org.joda.time.Period &  25.1\% &  45.1\% &  73.4\% &  82.2\% &  33.3\% &  50.0\% &  83.3\% &  100.0\%\\ 
Time-7 &  org.joda.time.format.DateTimeFormatter &  39.6\% &  46.1\% &  54.6\% &  59.3\% &  33.3\% &  16.7\% &  16.7\% &  0.0\%\\ 
Time-8 &  org.joda.time.DateTimeZone &  33.8\% &  47.6\% &  51.6\% &  53.8\% &  66.7\% &  83.3\% &  66.7\% &  100.0\%\\ 
\midrule 
Average &  &  23.4\% &  30.6\% &  37.7\% &  41.0\% &  21.8\% &  24.3\% &  31.1\% &  33.3\%\\ 
\bottomrule 
\end{tabular}