### 2017-02-18 papers/sbst17/mainTable.tex
Update scripts and add some tables

\begin{tabular}{ l rrrrrrr rrrrrrrr}\toprule 
 Benchmark &  \multicolumn{7}{c}{Branch Coverage} &  \multicolumn{7}{c}{Mutation Score}\\ 
 & 10s & 30s & 60s & 120s & 240s & 300s & 480s & 10s & 30s & 60s & 120s & 240s & 300s & 480s \\ 
\midrule 
BCEL-1 &  14.0\% &  0.0\% &  18.0\% &  36.6\% &  17.2\% &  0.0\% &  0.0\% &  5.9\% &  0.0\% &  14.9\% &  18.6\% &  16.8\% &  0.0\% &  0.0\%\\ 
BCEL-10 &  1.4\% &  1.4\% &  41.2\% &  45.8\% &  71.3\% &  76.4\% &  77.3\% &  1.3\% &  1.3\% &  35.6\% &  39.6\% &  62.2\% &  68.9\% &  65.3\%\\ 
BCEL-2 &  1.1\% &  1.3\% &  1.5\% &  1.9\% &  2.0\% &  2.4\% &  2.1\% &  0.4\% &  0.4\% &  0.7\% &  1.1\% &  1.2\% &  1.5\% &  1.5\%\\ 
BCEL-3 &  57.9\% &  63.8\% &  65.7\% &  73.3\% &  78.6\% &  79.8\% &  52.9\% &  33.9\% &  38.2\% &  29.4\% &  43.6\% &  44.4\% &  49.7\% &  53.6\%\\ 
BCEL-4 &  62.4\% &  66.2\% &  74.3\% &  80.2\% &  81.0\% &  82.6\% &  84.6\% &  46.3\% &  45.8\% &  58.2\% &  55.9\% &  59.9\% &  62.1\% &  81.9\%\\ 
BCEL-5 &  6.9\% &  6.9\% &  16.7\% &  55.1\% &  53.7\% &  31.9\% &  51.9\% &  4.8\% &  3.9\% &  17.4\% &  50.2\% &  50.7\% &  32.4\% &  48.3\%\\ 
BCEL-6 &  55.6\% &  54.9\% &  59.9\% &  66.7\% &  68.5\% &  68.5\% &  68.5\% &  56.7\% &  55.3\% &  66.0\% &  69.3\% &  76.7\% &  76.0\% &  78.0\%\\ 
BCEL-7 &  57.8\% &  63.7\% &  37.8\% &  63.7\% &  65.2\% &  21.5\% &  71.9\% &  38.1\% &  43.7\% &  27.8\% &  49.2\% &  65.1\% &  14.3\% &  50.8\%\\ 
BCEL-8 &  37.5\% &  37.5\% &  93.1\% &  93.1\% &  97.2\% &  93.1\% &  94.4\% &  21.9\% &  21.9\% &  91.7\% &  93.8\% &  94.8\% &  92.7\% &  94.8\%\\ 
BCEL-9 &  0.0\% &  0.0\% &  36.9\% &  0.0\% &  23.1\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  33.3\% &  0.0\% &  17.5\% &  0.0\% &  0.0\%\\ 
FREEHEP-1 &  83.6\% &  83.6\% &  87.4\% &  89.1\% &  92.5\% &  93.4\% &  93.4\% &  21.8\% &  23.3\% &  38.8\% &  37.4\% &  40.2\% &  38.2\% &  37.9\%\\ 
FREEHEP-10 &  48.6\% &  67.6\% &  82.4\% &  95.4\% &  94.0\% &  95.8\% &  95.8\% &  22.0\% &  22.9\% &  69.9\% &  77.4\% &  72.6\% &  78.3\% &  77.1\%\\ 
FREEHEP-2 &  0.0\% &  0.0\% &  64.2\% &  95.4\% &  96.1\% &  96.8\% &  96.5\% &  0.0\% &  0.0\% &  4.0\% &  7.1\% &  7.4\% &  7.6\% &  7.1\%\\ 
FREEHEP-3 &  0.0\% &  0.0\% &  48.1\% &  74.8\% &  86.7\% &  81.7\% &  87.1\% &  0.0\% &  0.0\% &  11.1\% &  14.4\% &  16.2\% &  19.2\% &  21.5\%\\ 
FREEHEP-4 &  13.2\% &  50.8\% &  49.2\% &  62.4\% &  79.4\% &  70.9\% &  84.7\% &  4.2\% &  18.0\% &  17.5\% &  24.7\% &  31.7\% &  28.9\% &  34.4\%\\ 
FREEHEP-5 &  32.1\% &  32.7\% &  59.1\% &  68.5\% &  67.3\% &  72.4\% &  75.2\% &  17.8\% &  21.5\% &  46.2\% &  52.3\% &  53.3\% &  53.3\% &  56.3\%\\ 
FREEHEP-6 &  0.0\% &  0.0\% &  86.4\% &  87.7\% &  88.3\% &  89.5\% &  89.5\% &  0.0\% &  0.0\% &  38.0\% &  46.7\% &  50.0\% &  41.3\% &  46.0\%\\ 
FREEHEP-7 &  54.2\% &  55.6\% &  85.4\% &  93.1\% &  94.4\% &  99.3\% &  93.8\% &  14.2\% &  14.2\% &  42.8\% &  51.2\% &  61.4\% &  47.5\% &  45.9\%\\ 
FREEHEP-8 &  25.9\% &  28.2\% &  58.8\% &  82.4\% &  81.0\% &  86.1\% &  91.2\% &  1.0\% &  2.4\% &  10.2\% &  51.8\% &  54.3\% &  43.1\% &  65.9\%\\ 
FREEHEP-9 &  1.4\% &  1.4\% &  13.1\% &  22.5\% &  49.1\% &  29.7\% &  33.8\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  16.7\% &  15.8\% &  16.2\%\\ 
GSON-1 &  0.0\% &  0.0\% &  46.3\% &  68.5\% &  70.4\% &  70.4\% &  70.4\% &  0.0\% &  0.0\% &  45.4\% &  63.0\% &  65.7\% &  65.7\% &  63.9\%\\ 
GSON-10 &  43.9\% &  45.1\% &  63.0\% &  62.2\% &  61.8\% &  62.2\% &  62.2\% &  22.9\% &  22.1\% &  54.5\% &  53.2\% &  52.4\% &  53.7\% &  55.4\%\\ 
GSON-2 &  3.3\% &  9.6\% &  25.5\% &  32.4\% &  36.7\% &  37.1\% &  39.6\% &  2.8\% &  8.1\% &  26.4\% &  34.1\% &  36.4\% &  37.2\% &  42.1\%\\ 
GSON-3 &  60.8\% &  54.6\% &  76.7\% &  80.4\% &  82.9\% &  85.4\% &  88.3\% &  46.7\% &  47.9\% &  78.9\% &  78.2\% &  82.4\% &  83.5\% &  77.8\%\\ 
GSON-4 &  3.1\% &  2.4\% &  13.1\% &  18.9\% &  27.5\% &  41.7\% &  46.8\% &  0.4\% &  0.6\% &  8.7\% &  12.8\% &  19.5\% &  17.2\% &  34.2\%\\ 
GSON-5 &  16.0\% &  21.8\% &  30.7\% &  32.7\% &  38.7\% &  40.4\% &  50.7\% &  17.7\% &  22.0\% &  32.8\% &  33.6\% &  37.7\% &  40.3\% &  44.3\%\\ 
GSON-6 &  17.4\% &  24.4\% &  63.6\% &  86.4\% &  86.4\% &  91.9\% &  91.9\% &  15.3\% &  19.6\% &  51.9\% &  72.0\% &  61.8\% &  74.2\% &  76.6\%\\ 
GSON-7 &  42.1\% &  43.0\% &  78.1\% &  82.5\% &  83.3\% &  82.5\% &  81.6\% &  71.3\% &  72.0\% &  90.0\% &  92.0\% &  92.7\% &  91.3\% &  91.3\%\\ 
GSON-9 &  32.8\% &  40.0\% &  51.1\% &  53.9\% &  46.7\% &  40.0\% &  48.3\% &  17.4\% &  27.4\% &  63.7\% &  64.7\% &  58.7\% &  53.7\% &  60.2\%\\ 
IMAGE-1 &  21.3\% &  27.8\% &  44.6\% &  55.4\% &  53.3\% &  49.8\% &  57.0\% &  7.5\% &  19.9\% &  18.4\% &  21.5\% &  28.0\% &  32.1\% &  36.4\%\\ 
IMAGE-2 &  65.0\% &  70.0\% &  78.9\% &  88.3\% &  87.8\% &  91.7\% &  96.1\% &  21.2\% &  20.8\% &  56.6\% &  62.8\% &  72.2\% &  72.6\% &  72.9\%\\ 
IMAGE-3 &  27.8\% &  29.0\% &  24.9\% &  28.5\% &  26.8\% &  39.0\% &  44.4\% &  8.6\% &  8.6\% &  9.2\% &  9.7\% &  14.4\% &  18.1\% &  21.4\%\\ 
IMAGE-4 &  33.6\% &  35.8\% &  47.3\% &  75.9\% &  79.4\% &  77.9\% &  77.9\% &  16.7\% &  17.4\% &  38.0\% &  63.4\% &  55.4\% &  64.1\% &  71.4\%\\ 
JXPATH-1 &  62.4\% &  65.4\% &  73.5\% &  76.8\% &  77.5\% &  76.8\% &  77.4\% &  41.3\% &  38.5\% &  50.3\% &  68.0\% &  68.3\% &  66.7\% &  67.2\%\\ 
JXPATH-10 &  17.6\% &  23.9\% &  35.5\% &  22.3\% &  52.2\% &  48.2\% &  48.0\% &  12.7\% &  15.4\% &  24.2\% &  19.0\% &  39.6\% &  35.4\% &  42.6\%\\ 
JXPATH-2 &  59.7\% &  67.1\% &  69.0\% &  74.4\% &  70.9\% &  77.9\% &  73.6\% &  40.9\% &  44.3\% &  50.8\% &  65.2\% &  62.5\% &  68.2\% &  62.9\%\\ 
JXPATH-3 &  73.4\% &  83.3\% &  79.7\% &  82.8\% &  83.3\% &  87.0\% &  90.1\% &  28.6\% &  39.5\% &  70.1\% &  86.4\% &  84.4\% &  93.2\% &  89.1\%\\ 
JXPATH-4 &  76.8\% &  78.1\% &  80.1\% &  83.0\% &  85.3\% &  89.5\% &  87.6\% &  68.5\% &  68.1\% &  71.4\% &  75.0\% &  78.3\% &  82.6\% &  83.7\%\\ 
JXPATH-5 &  66.0\% &  73.3\% &  82.7\% &  86.7\% &  88.0\% &  88.7\% &  93.3\% &  31.0\% &  31.0\% &  75.4\% &  89.7\% &  88.1\% &  90.5\% &  95.2\%\\ 
JXPATH-6 &  0.0\% &  0.0\% &  0.0\% &  48.6\% &  84.8\% &  83.3\% &  83.3\% &  0.0\% &  0.0\% &  0.0\% &  51.3\% &  82.7\% &  86.7\% &  78.7\%\\ 
JXPATH-7 &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
JXPATH-8 &  38.9\% &  40.2\% &  52.0\% &  52.4\% &  63.6\% &  60.4\% &  67.3\% &  22.1\% &  21.3\% &  39.7\% &  42.8\% &  53.5\% &  49.6\% &  56.6\%\\ 
JXPATH-9 &  9.5\% &  11.9\% &  45.9\% &  91.2\% &  85.4\% &  88.1\% &  91.8\% &  1.1\% &  3.3\% &  26.8\% &  59.8\% &  55.7\% &  54.4\% &  63.1\%\\ 
LA4J-1 &  36.2\% &  68.5\% &  73.5\% &  77.2\% &  77.5\% &  87.0\% &  79.9\% &  23.9\% &  47.1\% &  55.6\% &  55.9\% &  64.7\% &  74.5\% &  55.9\%\\ 
LA4J-10 &  31.8\% &  83.3\% &  90.9\% &  92.4\% &  92.4\% &  92.4\% &  92.4\% &  5.7\% &  22.6\% &  49.1\% &  70.4\% &  77.4\% &  69.2\% &  69.2\%\\ 
LA4J-2 &  10.7\% &  19.4\% &  19.8\% &  79.1\% &  81.4\% &  76.0\% &  78.1\% &  1.0\% &  2.2\% &  4.7\% &  14.8\% &  10.2\% &  17.0\% &  17.8\%\\ 
LA4J-3 &  0.0\% &  15.6\% &  0.0\% &  17.1\% &  38.9\% &  0.0\% &  20.3\% &  0.0\% &  10.8\% &  0.0\% &  10.5\% &  27.1\% &  0.0\% &  7.1\%\\ 
LA4J-4 &  0.0\% &  49.5\% &  71.7\% &  68.2\% &  71.7\% &  77.8\% &  80.8\% &  0.0\% &  33.1\% &  48.9\% &  52.3\% &  53.0\% &  43.9\% &  58.4\%\\ 
LA4J-5 &  0.0\% &  23.5\% &  32.6\% &  42.3\% &  54.4\% &  52.0\% &  44.4\% &  0.0\% &  14.9\% &  20.5\% &  24.1\% &  34.1\% &  20.3\% &  27.2\%\\ 
LA4J-6 &  13.3\% &  61.7\% &  70.0\% &  93.3\% &  95.0\% &  93.3\% &  91.7\% &  4.6\% &  16.7\% &  26.9\% &  78.7\% &  67.6\% &  63.0\% &  67.6\%\\ 
LA4J-7 &  0.0\% &  7.5\% &  0.0\% &  21.7\% &  0.0\% &  0.0\% &  21.9\% &  0.0\% &  6.1\% &  0.0\% &  17.6\% &  0.0\% &  0.0\% &  17.1\%\\ 
LA4J-8 &  32.4\% &  59.8\% &  74.0\% &  75.5\% &  80.9\% &  77.0\% &  77.9\% &  34.2\% &  58.5\% &  69.1\% &  72.7\% &  82.7\% &  78.8\% &  75.5\%\\ 
LA4J-9 &  15.0\% &  39.8\% &  67.4\% &  78.5\% &  77.0\% &  84.8\% &  93.1\% &  0.4\% &  5.6\% &  16.6\% &  24.2\% &  28.0\% &  29.4\% &  22.8\%\\ 
OKHTTP-1 &  4.2\% &  6.4\% &  5.9\% &  8.7\% &  14.4\% &  17.8\% &  16.8\% &  2.2\% &  3.4\% &  4.9\% &  6.5\% &  10.5\% &  8.0\% &  10.5\%\\ 
OKHTTP-2 &  33.3\% &  33.3\% &  33.3\% &  50.0\% &  50.0\% &  33.3\% &  50.0\% &  36.3\% &  38.2\% &  32.4\% &  48.0\% &  49.0\% &  32.4\% &  50.0\%\\ 
OKHTTP-3 &  70.7\% &  75.3\% &  75.8\% &  77.3\% &  74.7\% &  76.8\% &  76.3\% &  36.2\% &  37.6\% &  67.6\% &  69.5\% &  68.1\% &  68.5\% &  68.1\%\\ 
OKHTTP-4 &  29.6\% &  33.9\% &  54.8\% &  53.8\% &  55.9\% &  56.5\% &  58.6\% &  14.1\% &  17.4\% &  46.9\% &  54.5\% &  55.4\% &  55.9\% &  55.9\%\\ 
OKHTTP-5 &  0.0\% &  0.0\% &  3.6\% &  11.1\% &  23.8\% &  12.7\% &  33.3\% &  0.0\% &  0.0\% &  5.0\% &  12.0\% &  19.0\% &  11.8\% &  29.3\%\\ 
OKHTTP-6 &  54.8\% &  72.9\% &  77.6\% &  76.2\% &  83.3\% &  83.3\% &  83.8\% &  9.6\% &  15.3\% &  79.7\% &  75.7\% &  80.2\% &  83.6\% &  81.9\%\\ 
OKHTTP-7 &  27.6\% &  36.5\% &  24.0\% &  39.6\% &  42.7\% &  26.0\% &  46.4\% &  10.1\% &  12.2\% &  11.6\% &  28.6\% &  34.4\% &  12.7\% &  28.0\%\\ 
OKHTTP-8 &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\% &  0.0\%\\ 
RE2J-1 &  30.4\% &  50.7\% &  49.2\% &  51.4\% &  60.0\% &  60.5\% &  65.3\% &  19.7\% &  33.2\% &  32.0\% &  34.3\% &  43.8\% &  42.9\% &  39.9\%\\ 
RE2J-2 &  9.2\% &  73.8\% &  77.7\% &  92.0\% &  94.0\% &  95.2\% &  96.4\% &  3.9\% &  32.4\% &  36.6\% &  67.3\% &  69.6\% &  65.7\% &  71.6\%\\ 
RE2J-3 &  1.8\% &  1.8\% &  36.9\% &  48.8\% &  48.8\% &  50.0\% &  57.7\% &  3.4\% &  3.4\% &  23.0\% &  41.4\% &  43.7\% &  43.7\% &  48.9\%\\ 
RE2J-4 &  81.2\% &  85.1\% &  93.1\% &  94.8\% &  95.1\% &  95.5\% &  96.9\% &  27.6\% &  27.1\% &  91.5\% &  91.2\% &  92.0\% &  89.5\% &  84.6\%\\ 
RE2J-5 &  11.0\% &  22.7\% &  57.3\% &  61.2\% &  65.9\% &  58.0\% &  80.4\% &  2.9\% &  3.9\% &  31.1\% &  34.6\% &  36.9\% &  29.4\% &  42.7\%\\ 
RE2J-6 &  0.0\% &  0.0\% &  55.1\% &  56.7\% &  83.7\% &  84.3\% &  86.8\% &  0.0\% &  0.0\% &  42.0\% &  51.0\% &  69.2\% &  68.9\% &  66.3\%\\ 
RE2J-7 &  15.4\% &  39.2\% &  40.3\% &  53.2\% &  48.2\% &  31.4\% &  51.5\% &  5.9\% &  17.0\% &  15.4\% &  40.1\% &  39.2\% &  27.5\% &  47.8\%\\ 
RE2J-8 &  59.2\% &  90.8\% &  89.5\% &  92.9\% &  91.5\% &  92.9\% &  93.2\% &  26.7\% &  44.1\% &  67.7\% &  75.6\% &  81.0\% &  75.6\% &  81.2\%\\ 
\midrule 
Average  &  27.4\% &  36.4\% &  50.9\% &  60.6\% &  64.6\% &  62.3\% &  66.5\% &  15.3\% &  19.8\% &  36.5\% &  45.7\% &  49.6\% &  46.6\% &  50.7\%\\ 
\bottomrule 
\end{tabular}