### 2017-02-20 papers/sbst17/mainTable.tex
overall results

\begin{tabular}{ l rrrrrrr rrrrrrrr}\toprule 
 Benchmark &  \multicolumn{7}{c}{Branch Coverage} &  \multicolumn{7}{c}{Mutation Score}\\ 
 & 10s & 30s & 60s & 120s & 240s & 300s & 480s & 10s & 30s & 60s & 120s & 240s & 300s & 480s \\ 
\midrule 
BCEL-1 & 14.0\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 18.0\% & 36.6\% & 17.2\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 5.9\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 14.9\% & 18.6\% & 16.8\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\%\\ 
BCEL-10 & 1.4\% & 1.4\% & 41.2\% & 45.8\% & 71.3\% & 76.4\% & 77.3\% & 1.3\% & 1.3\% & 35.6\% & 39.6\% & 62.2\% & 68.9\% & 65.3\%\\ 
BCEL-2 & 1.1\% & 1.3\% & 1.5\% & 1.9\% & 2.0\% & 2.4\% & 2.1\% & 0.4\% & 0.4\% & 0.7\% & 1.1\% & 1.2\% & 1.5\% & 1.5\%\\ 
BCEL-3 & 57.9\% & 63.8\% & 65.7\% & 73.3\% & 78.6\% & 79.8\% & 52.9\% & 33.9\% & 38.2\% & 29.4\% & 43.6\% & 44.4\% & 49.7\% & 53.6\%\\ 
BCEL-4 & 62.4\% & 66.2\% & 74.3\% & 80.2\% & 81.0\% & 82.6\% & 84.6\% & 46.3\% & 45.8\% & 58.2\% & 55.9\% & 59.9\% & 62.1\% & 81.9\%\\ 
BCEL-5 & 6.9\% & 6.9\% & 16.7\% & 55.1\% & 53.7\% & 31.9\% & 51.9\% & 4.8\% & 3.9\% & 17.4\% & 50.2\% & 50.7\% & 32.4\% & 48.3\%\\ 
BCEL-6 & 55.6\% & 54.9\% & 59.9\% & 66.7\% & 68.5\% & 68.5\% & 68.5\% & 56.7\% & 55.3\% & 66.0\% & 69.3\% & 76.7\% & 76.0\% & 78.0\%\\ 
BCEL-7 & 57.8\% & 63.7\% & 37.8\% & 63.7\% & 65.2\% & 21.5\% & 71.9\% & 38.1\% & 43.7\% & 27.8\% & 49.2\% & 65.1\% & 14.3\% & 50.8\%\\ 
BCEL-8 & 37.5\% & 37.5\% & 93.1\% & 93.1\% & 97.2\% & 93.1\% & 94.4\% & 21.9\% & 21.9\% & 91.7\% & 93.8\% & 94.8\% & 92.7\% & 94.8\%\\ 
BCEL-9 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 36.9\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 23.1\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 33.3\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 17.5\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\%\\ 
FREEHEP-1 & 83.6\% & 83.6\% & 87.4\% & 89.1\% & 92.5\% & 93.4\% & 93.4\% & 21.8\% & 23.3\% & 38.8\% & 37.4\% & 40.2\% & 38.2\% & 37.9\%\\ 
FREEHEP-10 & 48.6\% & 67.6\% & 82.4\% & 95.4\% & 94.0\% & 95.8\% & 95.8\% & 22.0\% & 22.9\% & 69.9\% & 77.4\% & 72.6\% & 78.3\% & 77.1\%\\ 
FREEHEP-2 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 64.2\% & 95.4\% & 96.1\% & 96.8\% & 96.5\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 4.0\% & 7.1\% & 7.4\% & 7.6\% & 7.1\%\\ 
FREEHEP-3 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 48.1\% & 74.8\% & 86.7\% & 81.7\% & 87.1\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 11.1\% & 14.4\% & 16.2\% & 19.2\% & 21.5\%\\ 
FREEHEP-4 & 13.2\% & 50.8\% & 49.2\% & 62.4\% & 79.4\% & 70.9\% & 84.7\% & 4.2\% & 18.0\% & 17.5\% & 24.7\% & 31.7\% & 28.9\% & 34.4\%\\ 
FREEHEP-5 & 32.1\% & 32.7\% & 59.1\% & 68.5\% & 67.3\% & 72.4\% & 75.2\% & 17.8\% & 21.5\% & 46.2\% & 52.3\% & 53.3\% & 53.3\% & 56.3\%\\ 
FREEHEP-6 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 86.4\% & 87.7\% & 88.3\% & 89.5\% & 89.5\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 38.0\% & 46.7\% & 50.0\% & 41.3\% & 46.0\%\\ 
FREEHEP-7 & 54.2\% & 55.6\% & 85.4\% & 93.1\% & 94.4\% & 99.3\% & 93.8\% & 14.2\% & 14.2\% & 42.8\% & 51.2\% & 61.4\% & 47.5\% & 45.9\%\\ 
FREEHEP-8 & 25.9\% & 28.2\% & 58.8\% & 82.4\% & 81.0\% & 86.1\% & 91.2\% & 1.0\% & 2.4\% & 10.2\% & 51.8\% & 54.3\% & 43.1\% & 65.9\%\\ 
FREEHEP-9 & 1.4\% & 1.4\% & 13.1\% & 22.5\% & 49.1\% & 29.7\% & 33.8\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 16.7\% & 15.8\% & 16.2\%\\ 
GSON-1 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 46.3\% & 68.5\% & 70.4\% & 70.4\% & 70.4\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 45.4\% & 63.0\% & 65.7\% & 65.7\% & 63.9\%\\ 
GSON-10 & 43.9\% & 45.1\% & 63.0\% & 62.2\% & 61.8\% & 62.2\% & 62.2\% & 22.9\% & 22.1\% & 54.5\% & 53.2\% & 52.4\% & 53.7\% & 55.4\%\\ 
GSON-2 & 3.3\% & 9.6\% & 25.5\% & 32.4\% & 36.7\% & 37.1\% & 39.6\% & 2.8\% & 8.1\% & 26.4\% & 34.1\% & 36.4\% & 37.2\% & 42.1\%\\ 
GSON-3 & 60.8\% & 54.6\% & 76.7\% & 80.4\% & 82.9\% & 85.4\% & 88.3\% & 46.7\% & 47.9\% & 78.9\% & 78.2\% & 82.4\% & 83.5\% & 77.8\%\\ 
GSON-4 & 3.1\% & 2.4\% & 13.1\% & 18.9\% & 27.5\% & 41.7\% & 46.8\% & 0.4\% & 0.6\% & 8.7\% & 12.8\% & 19.5\% & 17.2\% & 34.2\%\\ 
GSON-5 & 16.0\% & 21.8\% & 30.7\% & 32.7\% & 38.7\% & 40.4\% & 50.7\% & 17.7\% & 22.0\% & 32.8\% & 33.6\% & 37.7\% & 40.3\% & 44.3\%\\ 
GSON-6 & 17.4\% & 24.4\% & 63.6\% & 86.4\% & 86.4\% & 91.9\% & 91.9\% & 15.3\% & 19.6\% & 51.9\% & 72.0\% & 61.8\% & 74.2\% & 76.6\%\\ 
GSON-7 & 42.1\% & 43.0\% & 78.1\% & 82.5\% & 83.3\% & 82.5\% & 81.6\% & 71.3\% & 72.0\% & 90.0\% & 92.0\% & 92.7\% & 91.3\% & 91.3\%\\ 
GSON-9 & 32.8\% & 40.0\% & 51.1\% & 53.9\% & 46.7\% & 40.0\% & 48.3\% & 17.4\% & 27.4\% & 63.7\% & 64.7\% & 58.7\% & 53.7\% & 60.2\%\\ 
IMAGE-1 & 21.3\% & 27.8\% & 44.6\% & 55.4\% & 53.3\% & 49.8\% & 57.0\% & 7.5\% & 19.9\% & 18.4\% & 21.5\% & 28.0\% & 32.1\% & 36.4\%\\ 
IMAGE-2 & 65.0\% & 70.0\% & 78.9\% & 88.3\% & 87.8\% & 91.7\% & 96.1\% & 21.2\% & 20.8\% & 56.6\% & 62.8\% & 72.2\% & 72.6\% & 72.9\%\\ 
IMAGE-3 & 27.8\% & 29.0\% & 24.9\% & 28.5\% & 26.8\% & 39.0\% & 44.4\% & 8.6\% & 8.6\% & 9.2\% & 9.7\% & 14.4\% & 18.1\% & 21.4\%\\ 
IMAGE-4 & 33.6\% & 35.8\% & 47.3\% & 75.9\% & 79.4\% & 77.9\% & 77.9\% & 16.7\% & 17.4\% & 38.0\% & 63.4\% & 55.4\% & 64.1\% & 71.4\%\\ 
JXPATH-1 & 62.4\% & 65.4\% & 73.5\% & 76.8\% & 77.5\% & 76.8\% & 77.4\% & 41.3\% & 38.5\% & 50.3\% & 68.0\% & 68.3\% & 66.7\% & 67.2\%\\ 
JXPATH-10 & 17.6\% & 23.9\% & 35.5\% & 22.3\% & 52.2\% & 48.2\% & 48.0\% & 12.7\% & 15.4\% & 24.2\% & 19.0\% & 39.6\% & 35.4\% & 42.6\%\\ 
JXPATH-2 & 59.7\% & 67.1\% & 69.0\% & 74.4\% & 70.9\% & 77.9\% & 73.6\% & 40.9\% & 44.3\% & 50.8\% & 65.2\% & 62.5\% & 68.2\% & 62.9\%\\ 
JXPATH-3 & 73.4\% & 83.3\% & 79.7\% & 82.8\% & 83.3\% & 87.0\% & 90.1\% & 28.6\% & 39.5\% & 70.1\% & 86.4\% & 84.4\% & 93.2\% & 89.1\%\\ 
JXPATH-4 & 76.8\% & 78.1\% & 80.1\% & 83.0\% & 85.3\% & 89.5\% & 87.6\% & 68.5\% & 68.1\% & 71.4\% & 75.0\% & 78.3\% & 82.6\% & 83.7\%\\ 
JXPATH-5 & 66.0\% & 73.3\% & 82.7\% & 86.7\% & 88.0\% & 88.7\% & 93.3\% & 31.0\% & 31.0\% & 75.4\% & 89.7\% & 88.1\% & 90.5\% & 95.2\%\\ 
JXPATH-6 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 48.6\% & 84.8\% & 83.3\% & 83.3\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 51.3\% & 82.7\% & 86.7\% & 78.7\%\\ 
JXPATH-7 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\%\\ 
JXPATH-8 & 38.9\% & 40.2\% & 52.0\% & 52.4\% & 63.6\% & 60.4\% & 67.3\% & 22.1\% & 21.3\% & 39.7\% & 42.8\% & 53.5\% & 49.6\% & 56.6\%\\ 
JXPATH-9 & 9.5\% & 11.9\% & 45.9\% & 91.2\% & 85.4\% & 88.1\% & 91.8\% & 1.1\% & 3.3\% & 26.8\% & 59.8\% & 55.7\% & 54.4\% & 63.1\%\\ 
LA4J-1 & 36.2\% & 68.5\% & 73.5\% & 77.2\% & 77.5\% & 87.0\% & 79.9\% & 23.9\% & 47.1\% & 55.6\% & 55.9\% & 64.7\% & 74.5\% & 55.9\%\\ 
LA4J-10 & 31.8\% & 83.3\% & 90.9\% & 92.4\% & 92.4\% & 92.4\% & 92.4\% & 5.7\% & 22.6\% & 49.1\% & 70.4\% & 77.4\% & 69.2\% & 69.2\%\\ 
LA4J-2 & 10.7\% & 19.4\% & 19.8\% & 79.1\% & 81.4\% & 76.0\% & 78.1\% & 1.0\% & 2.2\% & 4.7\% & 14.8\% & 10.2\% & 17.0\% & 17.8\%\\ 
LA4J-3 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 15.6\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 17.1\% & 38.9\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 20.3\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 10.8\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 10.5\% & 27.1\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 7.1\%\\ 
LA4J-4 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 49.5\% & 71.7\% & 68.2\% & 71.7\% & 77.8\% & 80.8\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 33.1\% & 48.9\% & 52.3\% & 53.0\% & 43.9\% & 58.4\%\\ 
LA4J-5 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 23.5\% & 32.6\% & 42.3\% & 54.4\% & 52.0\% & 44.4\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 14.9\% & 20.5\% & 24.1\% & 34.1\% & 20.3\% & 27.2\%\\ 
LA4J-6 & 13.3\% & 61.7\% & 70.0\% & 93.3\% & 95.0\% & 93.3\% & 91.7\% & 4.6\% & 16.7\% & 26.9\% & 78.7\% & 67.6\% & 63.0\% & 67.6\%\\ 
LA4J-7 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 7.5\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 21.7\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 21.9\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 6.1\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 17.6\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 17.1\%\\ 
LA4J-8 & 32.4\% & 59.8\% & 74.0\% & 75.5\% & 80.9\% & 77.0\% & 77.9\% & 34.2\% & 58.5\% & 69.1\% & 72.7\% & 82.7\% & 78.8\% & 75.5\%\\ 
LA4J-9 & 15.0\% & 39.8\% & 67.4\% & 78.5\% & 77.0\% & 84.8\% & 93.1\% & 0.4\% & 5.6\% & 16.6\% & 24.2\% & 28.0\% & 29.4\% & 22.8\%\\ 
OKHTTP-1 & 4.2\% & 6.4\% & 5.9\% & 8.7\% & 14.4\% & 17.8\% & 16.8\% & 2.2\% & 3.4\% & 4.9\% & 6.5\% & 10.5\% & 8.0\% & 10.5\%\\ 
OKHTTP-2 & 33.3\% & 33.3\% & 33.3\% & 50.0\% & 50.0\% & 33.3\% & 50.0\% & 36.3\% & 38.2\% & 32.4\% & 48.0\% & 49.0\% & 32.4\% & 50.0\%\\ 
OKHTTP-3 & 70.7\% & 75.3\% & 75.8\% & 77.3\% & 74.7\% & 76.8\% & 76.3\% & 36.2\% & 37.6\% & 67.6\% & 69.5\% & 68.1\% & 68.5\% & 68.1\%\\ 
OKHTTP-4 & 29.6\% & 33.9\% & 54.8\% & 53.8\% & 55.9\% & 56.5\% & 58.6\% & 14.1\% & 17.4\% & 46.9\% & 54.5\% & 55.4\% & 55.9\% & 55.9\%\\ 
OKHTTP-5 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 3.6\% & 11.1\% & 23.8\% & 12.7\% & 33.3\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 5.0\% & 12.0\% & 19.0\% & 11.8\% & 29.3\%\\ 
OKHTTP-6 & 54.8\% & 72.9\% & 77.6\% & 76.2\% & 83.3\% & 83.3\% & 83.8\% & 9.6\% & 15.3\% & 79.7\% & 75.7\% & 80.2\% & 83.6\% & 81.9\%\\ 
OKHTTP-7 & 27.6\% & 36.5\% & 24.0\% & 39.6\% & 42.7\% & 26.0\% & 46.4\% & 10.1\% & 12.2\% & 11.6\% & 28.6\% & 34.4\% & 12.7\% & 28.0\%\\ 
OKHTTP-8 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\%\\ 
RE2J-1 & 30.4\% & 50.7\% & 49.2\% & 51.4\% & 60.0\% & 60.5\% & 65.3\% & 19.7\% & 33.2\% & 32.0\% & 34.3\% & 43.8\% & 42.9\% & 39.9\%\\ 
RE2J-2 & 9.2\% & 73.8\% & 77.7\% & 92.0\% & 94.0\% & 95.2\% & 96.4\% & 3.9\% & 32.4\% & 36.6\% & 67.3\% & 69.6\% & 65.7\% & 71.6\%\\ 
RE2J-3 & 1.8\% & 1.8\% & 36.9\% & 48.8\% & 48.8\% & 50.0\% & 57.7\% & 3.4\% & 3.4\% & 23.0\% & 41.4\% & 43.7\% & 43.7\% & 48.9\%\\ 
RE2J-4 & 81.2\% & 85.1\% & 93.1\% & 94.8\% & 95.1\% & 95.5\% & 96.9\% & 27.6\% & 27.1\% & 91.5\% & 91.2\% & 92.0\% & 89.5\% & 84.6\%\\ 
RE2J-5 & 11.0\% & 22.7\% & 57.3\% & 61.2\% & 65.9\% & 58.0\% & 80.4\% & 2.9\% & 3.9\% & 31.1\% & 34.6\% & 36.9\% & 29.4\% & 42.7\%\\ 
RE2J-6 & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 55.1\% & 56.7\% & 83.7\% & 84.3\% & 86.8\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & \cellcolor{light-gray} \textcolor{black}{0.0}\% & 42.0\% & 51.0\% & 69.2\% & 68.9\% & 66.3\%\\ 
RE2J-7 & 15.4\% & 39.2\% & 40.3\% & 53.2\% & 48.2\% & 31.4\% & 51.5\% & 5.9\% & 17.0\% & 15.4\% & 40.1\% & 39.2\% & 27.5\% & 47.8\%\\ 
RE2J-8 & 59.2\% & 90.8\% & 89.5\% & 92.9\% & 91.5\% & 92.9\% & 93.2\% & 26.7\% & 44.1\% & 67.7\% & 75.6\% & 81.0\% & 75.6\% & 81.2\%\\ 
\midrule 
Average  &  27.4\% &  36.4\% &  50.9\% &  60.6\% &  64.6\% &  62.3\% &  66.5\% &  15.3\% &  19.8\% &  36.5\% &  45.7\% &  49.6\% &  46.6\% &  50.7\%\\ 
\bottomrule 
\end{tabular}