### 2018-01-31 papers/sbst18/mainTable.tex
updated table

\begin{tabular}{ l rrrrrrr rrrrrrrr}\toprule 
 Benchmark &  \multicolumn{7}{c}{Branch Coverage} &  \multicolumn{7}{c}{Mutation Score}\\ 
 & 10s & 30s & 60s & 120s & 240s & 300s & 480s & 10s & 30s & 60s & 120s & 240s & 300s & 480s \\ 
\midrule 
BCEL-1 & 54.7\% & 47.9\%\\ 
BCEL-10 & 62.5\% & 56.0\%\\ 
BCEL-2 & 1.2\% & 0.8\% & \multicolumn{13}{l}{Resulting test suite's coverage: 34\% (average coverage for all fitness functions)}\\ 
BCEL-3 & \cellcolor{light-gray} \textcolor{black}{0.0}\% &
                                                           \cellcolor{light-gray} \textcolor{black}{0.0}\%
& \multicolumn{13}{l}{org.evosuite.shaded.org.mockito.exceptions.misusing.InvalidUseOfMatchersException: 
Misplaced or misused argument matcher detected here}\\ 
BCEL-4 & 80.8\% & 80.5\%\\ 
BCEL-5 & 37.5\% & 63.8\%\\ 
BCEL-6 & 59.3\% & 76.0\%\\ 
BCEL-7 & 60.0\% & 50.0\%\\ 
BCEL-8 & 95.8\% & 90.6\%\\ 
BCEL-9 & 81.1\% & 67.5\%\\ 
FREEHEP-1 & 93.1\% & 42.2\%\\ 
FREEHEP-10 & 95.8\% & 69.6\%\\ 
FREEHEP-2 & 76.6\% & 5.7\%\\ 
FREEHEP-3 & 55.6\% & 62.1\%\\ 
FREEHEP-4 & 46.8\% & 22.2\%\\ 
FREEHEP-5 & 70.9\% & 54.1\%\\ 
FREEHEP-6 & 59.3\% & 27.0\%\\ 
FREEHEP-7 & 81.2\% & 77.2\%\\ 
FREEHEP-8 & 93.1\% & 74.7\%\\ 
FREEHEP-9 & 29.7\% & \cellcolor{light-gray} \textcolor{black}{0.0}\%
& \multicolumn{13}{l}{Resulting test suite's coverage: 36\%; ERROR ClassStateSupport - Could not initialize org.freehep.math.minuit.ModularFunctionMinimizer}\\ 
GSON-1 & 63.9\% & 41.7\%\\ 
GSON-10 & 58.5\% & 39.0\%\\ 
GSON-2 & 33.5\% & 34.8\%\\ 
GSON-3 & 77.5\% & 81.6\%\\ 
GSON-4 & 31.4\% & 24.2\%\\ 
GSON-5 & 29.3\% & 34.8\%\\ 
GSON-6 & 86.0\% & 72.6\%\\ 
GSON-7 & \cellcolor{light-gray} \textcolor{black}{0.0}\% &
                                                           \cellcolor{light-gray}
                                                           \textcolor{black}{0.0}\%
 & \multicolumn{13}{l}{java.lang.RuntimeException: not implemented: class sun.reflect.generics.reflectiveObjects.TypeVariableImpl}\\ 
GSON-8 & 86.0\% & 70.2\%\\ 
GSON-9 & \cellcolor{light-gray} \textcolor{black}{0.0}\% &
                                                           \cellcolor{light-gray} \textcolor{black}{0.0}\%
 & \multicolumn{13}{l}{java.lang.RuntimeException: not implemented: class sun.reflect.generics.reflectiveObjects.TypeVariableImpl}\\ 
IMAGE-1 & \cellcolor{light-gray} \textcolor{black}{0.0}\% &
                                                            \cellcolor{light-gray}
                                                            \textcolor{black}{0.0}\%
 & \multicolumn{13}{l}{ERROR FunctionalMockStatement - Failed to use Mockito on class org.apache.commons.imaging.formats.tiff.write.TiffOutputDirectory: tried to access class org.apache.commons.imaging.formats.tiff.write.TiffOutputItem}\\ 
IMAGE-2 & 88.3\% & 99.0\%\\ 
IMAGE-3 & 57.0\% & 32.5\%\\ 
IMAGE-4 & 80.6\% & 73.9\%\\ 
JXPATH-1 & 27.5\% & 25.4\%\\ 
JXPATH-10 & 40.8\% & 33.5\%\\ 
JXPATH-2 & 82.6\% & 72.7\%\\ 
JXPATH-3 & 73.4\% & 83.7\%\\ 
JXPATH-4 & 78.4\% & 72.8\%\\ 
JXPATH-5 & 90.0\% & 92.9\%\\ 
JXPATH-6 & 93.5\% & 96.0\%\\ 
JXPATH-7 & 24.5\% & 10.8\%\\ 
JXPATH-8 & 52.7\% & 42.8\%\\ 
JXPATH-9 & 68.4\% & 27.9\%\\ 
LA4J-1 & 75.4\% & 81.4\%\\ 
LA4J-10 & 86.4\% & 67.9\%\\ 
LA4J-2 & \cellcolor{light-gray} \textcolor{black}{0.0}\% &
                                                           \cellcolor{light-gray}
                                                           \textcolor{black}{0.0}\%
                                                           & \multicolumn{13}{l}{ERROR FunctionalMockStatement - Failed to use Mockito on class org.la4j.iterator.RowMajorMatrixIterator}\\ 
LA4J-3 & 67.1\% & 37.6\%\\ 
LA4J-4 & 84.8\% & 74.7\%\\ 
LA4J-5 & 52.9\% & 34.5\%\\ 
LA4J-6 & \cellcolor{light-gray} \textcolor{black}{0.0}\% &
                                                           \cellcolor{light-gray}
                                                           \textcolor{black}{0.0}\%
 & \multicolumn{13}{l}{ERROR FunctionalMockStatement - Failed to use Mockito on class org.la4j.iterator.VectorIterator}\\ 
LA4J-7 & 64.3\% & 40.6\%\\ 
LA4J-8 & 76.5\% & 93.6\%\\ 
LA4J-9 & 94.3\% & 30.4\%\\ 
OKHTTP-1 & 7.2\% & 5.6\%\\ 
OKHTTP-2 & 50.0\% & 44.1\%\\ 
OKHTTP-3 & 74.2\% & 67.6\%\\ 
OKHTTP-4 & 58.1\% & 53.5\%\\ 
OKHTTP-5 & 3.0\% & 4.0\%\\ 
OKHTTP-6 & 77.1\% & 76.3\%\\ 
OKHTTP-7 & 46.9\% & 27.0\%\\ 
OKHTTP-8 & \cellcolor{light-gray} \textcolor{black}{0.0}\% &
                                                             \cellcolor{light-gray}
                                                             \textcolor{black}{0.0}\%
 & \multicolumn{13}{l}{Caused by: java.lang.ClassNotFoundException:
   Class okio/Buffer.class should be in target project, but could not be found!}\\ 
RE2J-1 & 54.5\% & 36.9\%\\ 
RE2J-2 & 93.8\% & 71.6\%\\ 
RE2J-3 & 50.0\% & 37.9\%\\ 
RE2J-4 & 97.9\% & 90.6\%\\ 
RE2J-5 & 57.6\% & 29.1\%\\ 
RE2J-6 & 83.5\% & 71.2\%\\ 
RE2J-7 & 47.1\% & 27.8\%\\ 
RE2J-8 & 91.8\% & 79.0\%\\ 
\midrule 
Average  &  57.9\% &  47.3\%\\ 
\bottomrule 
\end{tabular}