### 2019-03-16 papers/sbst19/mainTable.tex
SBST'19 init commit.

\begin{tabular}{ ll rrrr rrrr}\toprule 
 \multirow{2}{1in}{Benchmark} & \multirow{2}{1in}{Java Class} &  \multicolumn{4}{c}{Branch Coverage} &  \multicolumn{4}{c}{Mutation Score}\\\cmidrule(lr){3-6}\cmidrule(lr){7-10} 
 & & 10s & 60s & 120s & 240s & 10s & 60s & 120s & 240s \\ 
\midrule 
AUTHZFORCE-1  &  org.ow2.authzforce.core.pdp.impl.PdpBean & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 16.7\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 14.4\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-11  &  org.ow2.authzforce.core.pdp.impl.func.LogicalNOfFunction & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 4.8\% & 4.8\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-27  &  org.ow2.authzforce.core.pdp.impl.func.MapFunctionFactory & 75.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 75.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-32  &  org.ow2.authzforce.core.pdp.impl.func.SubstringFunction & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 7.7\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-33  &  org.ow2.authzforce.core.pdp.impl.SchemaHandler & 55.6\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 57.6\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-48  &  org.ow2.authzforce.core.pdp.impl.policy.PolicyVersions & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 12.1\% & 34.8\% & 60.6\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 16.7\% & 37.3\% & 69.6\%\\ 
AUTHZFORCE-5  &  org.ow2.authzforce.core.pdp.impl.CloseableAttributeProvider & 20.5\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 27.4\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-52  &  org.ow2.authzforce.core.pdp.impl.PepActionExpression & 91.7\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 100.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-63  &  org.ow2.authzforce.core.pdp.impl.combining.DPUnlessPDCombiningAlg & 10.0\% & 3.3\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 5.6\% & 1.9\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-65  &  org.ow2.authzforce.core.pdp.impl.combining.FirstApplicableCombiningAlg & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
DUBBO-2  &  com.alibaba.dubbo.common.utils.ReflectUtils & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FASTJSON-1  &  com.alibaba.fastjson.parser.JSONLexerBase & \cellcolor{light-gray} \textcolor{black}{-} & 23.0\% & 24.9\% & 31.6\% & \cellcolor{light-gray} \textcolor{black}{-} & 7.7\% & 8.8\% & 6.5\%\\ 
FASTJSON-3  &  com.alibaba.fastjson.parser.DefaultJSONParser & \cellcolor{light-gray} \textcolor{black}{-} & 13.4\% & 20.5\% & 25.7\% & \cellcolor{light-gray} \textcolor{black}{-} & 9.4\% & 14.1\% & 19.1\%\\ 
FESCAR-1  &  com.alibaba.fescar.core.protocol.transaction.BranchReportRequest & 97.9\% & 95.8\% & 96.9\% & 99.0\% & 97.7\% & 97.7\% & 98.5\% & 100.0\%\\ 
FESCAR-12  &  com.alibaba.fescar.core.rpc.netty.RpcServerHandler & 12.5\% & 87.5\% & 87.5\% & 87.5\% & 50.0\% & 100.0\% & 100.0\% & 100.0\%\\ 
FESCAR-18  &  com.alibaba.fescar.core.protocol.MergedWarpMessage & 61.9\% & 48.8\% & 56.0\% & 42.9\% & 81.2\% & 60.9\% & 68.1\% & 51.4\%\\ 
FESCAR-23  &  com.alibaba.fescar.core.protocol.MergeResultMessage & 63.1\% & 78.6\% & 64.3\% & 83.3\% & 78.6\% & 97.0\% & 79.2\% & 98.8\%\\ 
FESCAR-25  &  com.alibaba.fescar.core.rpc.netty.RmMessageListener & 37.5\% & 62.5\% & 62.5\% & 50.0\% & 22.2\% & 22.2\% & 22.2\% & 18.5\%\\ 
FESCAR-36  &  com.alibaba.fescar.core.protocol.RegisterRMRequest & 100.0\% & 92.9\% & 80.1\% & 83.3\% & 100.0\% & 93.8\% & 81.2\% & 83.0\%\\ 
FESCAR-37  &  com.alibaba.fescar.core.rpc.RpcContext & 52.9\% & 89.2\% & 91.2\% & 91.2\% & 71.9\% & 96.4\% & 94.8\% & 95.8\%\\ 
FESCAR-41  &  com.alibaba.fescar.core.rpc.netty.RmRpcClient & \cellcolor{light-gray} \textcolor{black}{-} & 2.0\% & 2.0\% & 2.0\% & \cellcolor{light-gray} \textcolor{black}{-} & 2.4\% & 2.4\% & 2.4\%\\ 
FESCAR-42  &  com.alibaba.fescar.core.rpc.DefaultServerMessageListenerImpl & \cellcolor{light-gray} \textcolor{black}{-} & 16.1\% & 27.4\% & 27.9\% & \cellcolor{light-gray} \textcolor{black}{-} & 16.7\% & 27.8\% & 29.2\%\\ 
FESCAR-7  &  com.alibaba.fescar.core.rpc.netty.MessageCodecHandler & 37.8\% & 78.4\% & 82.8\% & 87.1\% & 37.9\% & 92.8\% & 95.2\% & 98.3\%\\ 
OKIO-1  &  okio.Buffer & 27.4\% & 57.0\% & 62.3\% & 67.9\% & 13.2\% & 20.2\% & 16.3\% & 18.5\%\\ 
OKIO-4  &  okio.RealBufferedSource & 19.2\% & 37.6\% & 51.3\% & 76.2\% & 13.1\% & 28.8\% & 40.1\% & 45.3\%\\ 
SPOON-105  &  spoon.support.compiler.jdt.PositionBuilder & \cellcolor{light-gray} \textcolor{black}{-} & 22.1\% & 16.4\% & 3.8\% & \cellcolor{light-gray} \textcolor{black}{-} & 27.7\% & 23.1\% & 5.4\%\\ 
SPOON-155  &  spoon.reflect.visitor.filter.AllMethodsSameSignatureFunction & 12.0\% & 4.2\% & 2.5\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 19.0\% & 10.7\% & 4.3\% & 5.0\%\\ 
SPOON-16  &  spoon.reflect.path.CtElementPathBuilder & 34.0\% & 34.3\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 2.0\% & 45.5\% & 43.9\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 4.0\%\\ 
SPOON-169  &  spoon.reflect.visitor.ImportScannerImpl & \cellcolor{light-gray} \textcolor{black}{-} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{-} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-20  &  spoon.support.reflect.reference.CtLocalVariableReferenceImpl & 33.3\% & 20.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 40.3\% & 33.3\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-211  &  spoon.reflect.path.impl.CtRolePathElement & 8.1\% & 19.2\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 5.1\% & 27.8\% & 36.8\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 17.7\%\\ 
SPOON-25  &  spoon.pattern.internal.ValueConvertorImpl & \cellcolor{light-gray} \textcolor{black}{-} & \cellcolor{light-gray} \textcolor{black}{-} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{-} & \cellcolor{light-gray} \textcolor{black}{-} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-253  &  spoon.pattern.internal.parameter.MapParameterInfo & 29.7\% & 50.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 34.8\% & 50.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-32  &  spoon.MavenLauncher & 12.5\% & 10.4\% & 10.4\% & 12.5\% & 6.7\% & 5.6\% & 5.6\% & 6.7\%\\ 
SPOON-65  &  spoon.support.DefaultCoreFactory & 9.0\% & 49.8\% & 20.0\% & 5.7\% & 1.2\% & 2.3\% & 1.2\% & 2.0\%\\ 
WEBMAGIC-1  &  us.codecraft.webmagic.model.PageModelExtractor & 28.0\% & 27.6\% & 28.8\% & 31.3\% & 50.4\% & 50.0\% & 51.3\% & 54.3\%\\ 
WEBMAGIC-4  &  us.codecraft.webmagic.Page & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
ZXING-10  &  com.google.zxing.qrcode.encoder.Encoder & 70.1\% & 82.3\% & 82.9\% & 84.0\% & 63.0\% & 76.1\% & 72.6\% & 72.8\%\\ 
\midrule 
Average &  &  17.3\% &  17.2\% &  14.8\% &  15.5\% &  19.9\% &  16.9\% &  13.9\% &  14.6\%\\ 
\bottomrule 
\end{tabular}