### 2019-03-21 papers/sbst19/mainTable-t3.tex
T3 results.

\begin{tabular}{ ll rrrr rrrr}\toprule 
 \multirow{2}{1in}{Benchmark} & \multirow{2}{1in}{Java Class} &  \multicolumn{4}{c}{Branch Coverage} &  \multicolumn{4}{c}{Mutation Score}\\\cmidrule(lr){3-6}\cmidrule(lr){7-10} 
 & & 10s & 60s & 120s & 240s & 10s & 60s & 120s & 240s \\ 
\midrule 
AUTHZFORCE-1  &  org.ow2.authzforce.core.pdp.impl.PdpBean & 100.0\% & 100.0\% & 100.0\% & 100.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-11  &  org.ow2.authzforce.core.pdp.impl.func.LogicalNOfFunction & 10.0\% & 10.0\% & 10.0\% & 10.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-27  &  org.ow2.authzforce.core.pdp.impl.func.MapFunctionFactory & 25.0\% & 25.0\% & 25.0\% & 25.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-32  &  org.ow2.authzforce.core.pdp.impl.func.SubstringFunction & 20.0\% & 20.0\% & 20.0\% & 20.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-33  &  org.ow2.authzforce.core.pdp.impl.SchemaHandler & 94.4\% & 100.0\% & 100.0\% & 100.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-48  &  org.ow2.authzforce.core.pdp.impl.policy.PolicyVersions & 62.9\% & 58.3\% & 63.6\% & 58.3\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-5  &  org.ow2.authzforce.core.pdp.impl.CloseableAttributeProvider & 53.8\% & 54.5\% & 54.5\% & 54.5\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-52  &  org.ow2.authzforce.core.pdp.impl.PepActionExpression & 83.3\% & 83.3\% & 83.3\% & 83.3\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-63  &  org.ow2.authzforce.core.pdp.impl.combining.DPUnlessPDCombiningAlg & 30.0\% & 30.0\% & 30.0\% & 30.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
AUTHZFORCE-65  &  org.ow2.authzforce.core.pdp.impl.combining.FirstApplicableCombiningAlg & 41.7\% & 41.7\% & 41.7\% & 41.7\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
DUBBO-2  &  com.alibaba.dubbo.common.utils.ReflectUtils & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FASTJSON-1  &  com.alibaba.fastjson.parser.JSONLexerBase & 0.6\% & 0.8\% & 2.3\% & 4.3\% & 0.4\% & 0.6\% & 1.6\% & 1.3\%\\ 
FASTJSON-3  &  com.alibaba.fastjson.parser.DefaultJSONParser & 5.0\% & 15.0\% & 16.0\% & 16.7\% & 6.4\% & 12.9\% & 12.9\% & 11.6\%\\ 
FESCAR-1  &  com.alibaba.fescar.core.protocol.transaction.BranchReportRequest & 60.4\% & 64.6\% & 61.5\% & 68.8\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FESCAR-12  &  com.alibaba.fescar.core.rpc.netty.RpcServerHandler & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FESCAR-18  &  com.alibaba.fescar.core.protocol.MergedWarpMessage & 35.7\% & 35.7\% & 35.7\% & 35.7\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FESCAR-23  &  com.alibaba.fescar.core.protocol.MergeResultMessage & 28.6\% & 28.6\% & 28.6\% & 28.6\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FESCAR-25  &  com.alibaba.fescar.core.rpc.netty.RmMessageListener & 37.5\% & 37.5\% & 37.5\% & 37.5\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FESCAR-36  &  com.alibaba.fescar.core.protocol.RegisterRMRequest & 14.7\% & 14.1\% & 12.8\% & 12.8\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FESCAR-37  &  com.alibaba.fescar.core.rpc.RpcContext & 35.3\% & 35.3\% & 35.3\% & 35.3\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FESCAR-41  &  com.alibaba.fescar.core.rpc.netty.RmRpcClient & \cellcolor{light-gray} \textcolor{black}{0.0\%} & 0.7\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FESCAR-42  &  com.alibaba.fescar.core.rpc.DefaultServerMessageListenerImpl & 7.1\% & 7.1\% & 7.1\% & 7.1\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
FESCAR-7  &  com.alibaba.fescar.core.rpc.netty.MessageCodecHandler & 18.8\% & 18.8\% & 18.8\% & 18.8\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
OKIO-1  &  okio.Buffer & 22.9\% & 10.2\% & 18.3\% & 21.4\% & 8.5\% & 5.2\% & 1.6\% & 8.1\%\\ 
OKIO-4  &  okio.RealBufferedSource & 0.6\% & 0.6\% & 0.6\% & 0.6\% & 0.9\% & 0.9\% & 0.9\% & 0.9\%\\ 
SPOON-105  &  spoon.support.compiler.jdt.PositionBuilder & 17.3\% & 16.7\% & 16.7\% & 16.8\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-155  &  spoon.reflect.visitor.filter.AllMethodsSameSignatureFunction & 6.2\% & 6.2\% & 6.2\% & 6.2\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-16  &  spoon.reflect.path.CtElementPathBuilder & 7.8\% & 7.8\% & 7.8\% & 7.8\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-169  &  spoon.reflect.visitor.ImportScannerImpl & 3.3\% & 4.3\% & 5.2\% & 4.1\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-20  &  spoon.support.reflect.reference.CtLocalVariableReferenceImpl & 20.0\% & 20.0\% & 20.0\% & 20.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-211  &  spoon.reflect.path.impl.CtRolePathElement & 15.4\% & 15.4\% & 15.4\% & 15.4\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-25  &  spoon.pattern.internal.ValueConvertorImpl & 21.2\% & 20.8\% & 21.2\% & 21.2\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-253  &  spoon.pattern.internal.parameter.MapParameterInfo & 54.7\% & 52.6\% & 53.1\% & 52.1\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-32  &  spoon.MavenLauncher & 12.5\% & 12.5\% & 12.5\% & 12.5\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
SPOON-65  &  spoon.support.DefaultCoreFactory & 50.0\% & 50.0\% & 50.0\% & 50.0\% & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
WEBMAGIC-1  &  us.codecraft.webmagic.model.PageModelExtractor & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%} & \cellcolor{light-gray} \textcolor{black}{0.0\%}\\ 
WEBMAGIC-4  &  us.codecraft.webmagic.Page & 2.1\% & 6.3\% & 5.6\% & 5.6\% & 9.8\% & 22.4\% & 31.6\% & 14.9\%\\ 
ZXING-10  &  com.google.zxing.qrcode.encoder.Encoder & 62.0\% & 64.0\% & 63.9\% & 64.7\% & 33.8\% & 28.6\% & 34.1\% & 33.6\%\\ 
\midrule 
Average &  &  17.3\% &  17.2\% &  14.8\% &  15.5\% &  19.9\% &  16.9\% &  13.9\% &  14.6\%\\ 
\bottomrule 
\end{tabular}