### 2020-03-29 papers/sbst20/main.tex
Explaining more cases

\documentclass[sigconf]{acmart}
\usepackage{amsfonts}
\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage{listings}
\usepackage{algorithmic}
\usepackage{float}
%\usepackage[numbers,sort&compress,square]{natbib}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{subfigure}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{xcolor,colortbl}
%\usepackage[usenames,dvipsnames,]{xcolor}
%\usepackage{soul}
%\usepackage{boxedminipage}
%\usepackage{alltt}
\usepackage{multirow}
%\usepackage{paralist}
%\usepackage{amsmath}
\usepackage{balance}
\definecolor{light-gray}{gray}{0.90}

\newcommand{\EVOSUITE}{{\sc EvoSuite}\xspace}
\newcommand{\RANDOOP}{{\sc Randoop}\xspace}
\newcommand{\MUTEST}{{\sc $\mu$Test}\xspace}
\newcommand{\CS}{{\sc SF100}\xspace}
\newcommand{\TOTALPOINTS}{{255.43}\xspace}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[SBST '20]{SBST'20: IEEE/ACM 13th International Workshop on Search-Based Software Testing}{October 2020}{Seoul, South Korea}
\acmBooktitle{SBST'20: IEEE/ACM 13th International Workshop on Search-Based Software Testing,
  October 2020, Seoul, South Korea}
\acmPrice{15.00}
\acmISBN{978-1-4503-9999-9/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.

\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{\EVOSUITE at the SBST 2020 Tool Competition}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Annibale Panichella}
%\orcid{0000-0002-5385-7695}
%\author{G.K.M. Tobin}
%\authornotemark[1]
%\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Delft University of Technology}
  \state{The Netherlands}
}
\email{a.panichella@tudelft.nl}

\author{Jos\'e Campos}
%\email{a.panichella@tudelft.nl}
%\orcid{0000-0002-5385-7695}
%\author{G.K.M. Tobin}
%\authornotemark[1]
\affiliation{%
  \institution{LASIGE, Faculdade de Ci\^encias \\ Universidade de Lisboa}
  \city{Lisboa}
  \country{Portugal}
}
\email{jcmcampos@fc.ul.pt}

\author{Gordon Fraser}
\affiliation{%
  \institution{Chair of Software Engineering II, University of Passau}
  \city{Passau}
  \country{Germany}
}
\email{gordon.fraser@uni-passau.de}

%\begin{teaserfigure}

%\end{teaserfigure}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
%\renewcommand{\shortauthors}{Anonymous et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
TBD
\end{abstract}

\maketitle

\section{Introduction}

\section{\EVOSUITE}

\begin{table}[!h]
\renewcommand{\arraystretch}{1.3}
\caption{Classification of the \EVOSUITE unit test generation
  tool}\label{tool-description}
\resizebox{1.0\columnwidth}{!}{  
\begin{tabular}{|l|p{5cm}|}
  \hline
  \multicolumn{2}{|l|}{Prerequisites} \\
  \hline
  Static or dynamic &  Dynamic testing at the Java class level\\
  Software Type &  Java classes\\
  Lifecycle phase&  Unit testing for Java programs\\
  Environment&  All Java development environments \\
  Knowledge required & JUnit unit testing for Java\\
  Experience required &  Basic unit testing knowledge\\
 \hline
  \multicolumn{2}{|l|}{Input and Output of the tool} \\
  \hline
 Input & Bytecode of the target class and dependencies \\
\hline
Output&  JUnit 4 test cases\\

  \hline
  \multicolumn{2}{|l|}{Operation} \\
  \hline
  Interaction &  Through the command line, and plugins for IntelliJ, Maven and Eclipse\\
  User guidance &  Manual verification of assertions for functional faults\\
  Source of information &  http://www.evosuite.org \\
  Maturity&  Mature research prototype, under development\\
  Technology behind the tool & Search-based testing / many-objective optimization \\
\hline
  \multicolumn{2}{|l|}{Obtaining the tool and information} \\
  \hline
License & Lesser GPL V.3\\
Cost & Open source\\
Support & None \\
\hline
\hline
  \multicolumn{2}{|l|}{Does there exist empirical evidence about} \\
  \hline
  Effectiveness and Scalability & See~\cite{GoA_TSE12,fraser2014large} \\
%Completeness & \\
%Effectiveness & \\
%Efficiency & \\
%Defect types & \\
%Scalability & \\
%Comprehensibility & \\
%Learnability & \\
%Subjective satisfaction & \\
%Other & \\
\hline
\end{tabular}\vspace{-1em}
}
\end{table}

\section{Tool Setup}

\section{Benchmark Results}

\begin{table*}[t]
  \centering
  \caption{\label{table:results}Detailed results of \EVOSUITE on the
    SBST benchmark classes.}
\vspace{-1em}
\resizebox{0.8\textwidth}{!}{  
\input{mainTable.tex}
}	
\end{table*}

%% FESCAR-15 0% coverage 0% mutation score

\subsubsection*{FESCAR-15} According to the \texttt{transcript.csv} file, it
seems that the generated test cases do cover some lines of code but no
branches/conditions.  
%Looking at the \texttt{logs.log}, \EVOSUITE also reported
%0\% coverage and we could not find any issues in any %log file. 
For further investigation, we ran \EVOSUITE as stand-alone tool (i.e., wot using the benchmark infrastructure but it was still not capable of covering any line. Through manual investigation, we notice that the CUT manages objects of the class \texttt{ScheduledExecutorService} and \textit{threads}. Generating tests for CUTs that produce or manage threads is challenging and a well-known open problem in the literature~\cite{FrA13a}.
\textcolor{red}{Annibale: does EvoSuite allow to generate threads within the scope of the generated tests? If I am not mistaken, it does not for security reasons. Should we mention it?
}

%% FESCAR-2 0% coverage 0% mutation score

\subsubsection*{FESCAR-2} This CUT contains only a few branches (i.e., 14) and lines (i.e., 19). \EVOSUITE could generate test cases that could cover around 20\%-25\% of lines but 0\% of condition coverage. The generated tests cover branchless methods and the root branch of the method \texttt{lookup}, which is the only method with some conditions. It is worth noticing that \EVOSUITE could not complete a single generation, even with a search budget of 180s. The CUT is particularly expensive, and each generated test requires seconds for its execution. Low search budgets are thus not feasible nor recommended for this CUT.

\subsubsection*{FESCAR-41, FESCAR-6} \EVOSUITE reached a very low line and branch coverage for these two CUTs, independently from the search budget. These CUT also manage threads, and more precisely objects of the classes \texttt{ThreadPoolExecutor} and \texttt{NamedThreadFactory}.

%% FESCAR-37 86.8% coverage 0% mutation score
%% Annibale: from the logs, I see a \texttt{java.util.concurrent.ExecutionException}.
%% \subsubsection*{FESCAR-37} The classpath is incomplete as there is a
%% \texttt{java.lang.}-\texttt{ClassNotFoundException:  io.netty.channel.Channel}.

%% GUAVA-110 0% coverage 0.6% mutation score

\subsubsection*{GUAVA-110, GUAVA-240} Both classes make use of the annotation \texttt{@Nullable}. GUAVA-110 uses the annotation for the input parameter of the method \texttt{equals(...)}, while GUAVA-240 uses it for the methods \texttt{contains(...)} and \texttt{remove(...)}. The static analyzer implemented in \EVOSUITE does not handle the annotation \texttt{@Nullable} and triggers the error \textit{``Cannot find symbol''}.

\subsubsection*{GUAVA-47} For this class, \EVOSUITE reached very low coverage (both line and branch) and mutation score. To have a better understanding of the underlying issue, we ran \EVOSUITE stand-alone. We notice that our tool could not generate any single test cases, never completing the initialization process (and the generation of the initial population, in particular) even with a search budget of 180s. A more in-depth analysis is needed to discover the root cause of the issue.

\subsubsection*{PDFBOX-157} \EVOSUITE crashed in 9 out ten times with a search budget of 60s, and in 10 out of 10 runs for larger budgets.

%% FESCAR-7 73.3% coverage 0% mutation score
%% GUAVA-212 77.6% coverage 0% mutation score
%% GUAVA-22 51.8% coverage 0% mutation score
%% PDFBOX-265 51.8% coverage 0% mutation score

\subsubsection*{FESCAR-23, FESCAR-37, FESCAR-7, FESCAR-8, GUAVA-212, GUAVA-22, PDFBOX-265, PDFBOX-278, PDFBOX-285, PDFBOX-40, PDFBOX-91, PDFBOX-62, SPOON-105, SPOON-253} The mutation analysis
failed due to a \texttt{java.util.concurrent.ExecutionException} thrown by the
experimental infrastructure.

%% PDFBOX-235 0% coverage 0% mutation score
%% SPOON-155 0% coverage 0% mutation score

\subsubsection*{PDFBOX-235, SPOON-155} \EVOSUITE crashed due to some Mocking stuff.

\section{Conclusions}

\bibliographystyle{ACM-Reference-Format}
\bibliography{papers}

\end{document}