### 2020-04-04 papers/sbst20/main.tex
My pass on the paper

\documentclass[sigconf]{acmart}
\usepackage{amsfonts}
\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage{listings}
\usepackage{algorithmic}
\usepackage{float}
%\usepackage[numbers,sort&compress,square]{natbib}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{subfigure}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{xcolor,colortbl}
%\usepackage[usenames,dvipsnames,]{xcolor}
%\usepackage{soul}
%\usepackage{boxedminipage}
%\usepackage{alltt}
\usepackage{multirow}
%\usepackage{paralist}
%\usepackage{amsmath}
\usepackage{balance}
\definecolor{light-gray}{gray}{0.90}

\newcommand{\EVOSUITE}{{\sc EvoSuite}\xspace}
\newcommand{\RANDOOP}{{\sc Randoop}\xspace}
\newcommand{\MUTEST}{{\sc $\mu$Test}\xspace}
\newcommand{\CS}{{\sc SF100}\xspace}
\newcommand{\TOTALPOINTS}{{406.14}\xspace}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[SBST '20]{SBST'20: IEEE/ACM 13th International Workshop on Search-Based Software Testing}{October 2020}{Seoul, South Korea}
\acmBooktitle{SBST'20: IEEE/ACM 13th International Workshop on Search-Based Software Testing,
  October 2020, Seoul, South Korea}
\acmPrice{15.00}
\acmISBN{978-1-4503-9999-9/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.

\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{\EVOSUITE at the SBST 2020 Tool Competition}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Annibale Panichella}
%\orcid{0000-0002-5385-7695}
%\author{G.K.M. Tobin}
%\authornotemark[1]
%\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Delft University of Technology}
  \state{The Netherlands}
}
\email{a.panichella@tudelft.nl}

\author{Jos\'e Campos}
%\email{a.panichella@tudelft.nl}
%\orcid{0000-0002-5385-7695}
%\author{G.K.M. Tobin}
%\authornotemark[1]
\affiliation{%
  \institution{LASIGE, Faculdade de Ci\^encias \\ Universidade de Lisboa}
  \city{Lisboa}
  \country{Portugal}
}
\email{jcmcampos@fc.ul.pt}

\author{Gordon Fraser}
\affiliation{%
  \institution{Chair of Software Engineering II, University of Passau}
  \city{Passau}
  \country{Germany}
}
\email{gordon.fraser@uni-passau.de}

%\begin{teaserfigure}

%\end{teaserfigure}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
%\renewcommand{\shortauthors}{Anonymous et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
%
  \EVOSUITE is a search-based tool that automatically generates
  executable unit tests for Java code (JUnit tests).  This paper
  summarizes the results and experiences of \EVOSUITE's participation
  at the eighth unit testing competition at SBST 2020, where \EVOSUITE
  achieved the highest overall score (\TOTALPOINTS points) for the
  seventh time in eight editions of the competition.
  %
\end{abstract}

\maketitle

\section{Introduction}

Automated unit test generation can support developers and testers, produce
regression test suites, and is an enabler for dynamic program analyses. The
annual unit test generation aims to foster research and development of
automated unit test generators. This paper describes the results obtained by
the \EVOSUITE test generation tool~\cite{FrA11c} in this competition. \EVOSUITE
uses meta-heuristic search to evolve unit test suites with high
coverage, and automatically produces regression oracles in the form of test
assertions. In the 8th instance of the competition at the International
Workshop on Search-Based Software Testing (SBST) 2020, \EVOSUITE achieved an
overall score of \TOTALPOINTS, which was the highest among the competing and
baseline tools. This paper describes the results obtained by the \EVOSUITE test
generation tool in this competition.

\section{\EVOSUITE}

\begin{table}[!h]
\renewcommand{\arraystretch}{1.3}
\caption{Classification of the \EVOSUITE unit test generation
  tool}\label{tool-description}
\resizebox{1.0\columnwidth}{!}{  
\begin{tabular}{|l|p{5cm}|}
  \hline
  \multicolumn{2}{|l|}{Prerequisites} \\
  \hline
  Static or dynamic &  Dynamic testing at the Java class level\\
  Software Type &  Java classes\\
  Lifecycle phase&  Unit testing for Java programs\\
  Environment&  All Java development environments \\
  Knowledge required & JUnit unit testing for Java\\
  Experience required &  Basic unit testing knowledge\\
 \hline
  \multicolumn{2}{|l|}{Input and Output of the tool} \\
  \hline
 Input & Bytecode of the target class and dependencies \\
\hline
Output&  JUnit 4 test cases\\

  \hline
  \multicolumn{2}{|l|}{Operation} \\
  \hline
  Interaction &  Through the command line, and plugins for IntelliJ, Maven and Eclipse\\
  User guidance &  Manual verification of assertions for functional faults\\
  Source of information &  http://www.evosuite.org \\
  Maturity&  Mature research prototype, under development\\
  Technology behind the tool & Search-based testing / many-objective optimization \\
\hline
  \multicolumn{2}{|l|}{Obtaining the tool and information} \\
  \hline
License & Lesser GPL V.3\\
Cost & Open source\\
Support & None \\
\hline
\hline
  \multicolumn{2}{|l|}{Does there exist empirical evidence about} \\
  \hline
  Effectiveness and Scalability & See~\cite{GoA_TSE12,fraser2014large} \\
%Completeness & \\
%Effectiveness & \\
%Efficiency & \\
%Defect types & \\
%Scalability & \\
%Comprehensibility & \\
%Learnability & \\
%Subjective satisfaction & \\
%Other & \\
\hline
\end{tabular}\vspace{-1em}
}
\end{table}

\EVOSUITE~\cite{FrA11c} is a search-based unit test generation
tool~\cite{GoA_TSE12}. Table~\ref{tool-description} summarizes the features of
\EVOSUITE in the standard format of the SBST unit testing competition: Given
just the Java classpath containing all compiled dependencies and the name of a
class under test, \EVOSUITE automatically generates a set of JUnit test cases
aimed at maximizing code coverage. \EVOSUITE can be used on the command line,
as a Maven plugin, or using plugins for the Eclipse and IntelliJ development
environments~\cite{ICST16_Tool}.

Throughout its development, several different search algorithms have
been evaluated, starting with a basic genetic algorithm. \EVOSUITE's encoding of test cases for the evolutionary search (i.e., its chromosomes) consists of variable-length sequences of Java statements (e.g., primitive statements and calls on the class under test). The usual search operators used in evolutionary search (e.g., selection, crossover, mutation) are adapted for this particular representation. 
%
In the original approach, individuals of the genetic algorithm were whole test
suites, with the optimization goal of finding a test suite that maximizes code
coverage. This was later improved by adding an archive of
solutions~\cite{emse_archive} to keep the search focused on uncovered goals,
iteratively discarding covered goals and storing the tests that covered them.

Recently, the \textit{Dynamic Many-Objective Sorting Algorithm} (DynaMOSA) search algorithm~\cite{dynamosa, panichella:ssbse2018, panichella2018large}  has been shown to be the most effective approach of all the algorithms evaluated so far. DynaMOSA is a many-objective algorithm, where individuals of the search population are test cases, rather than test suites, and the optimization is driven by a collection of individual fitness functions, one for each coverage objective (e.g., line or branch).

The fitness functions in \EVOSUITE are based on traditional heuristics for code
coverage, such as the branch distance and the approach level
(see~\cite{GoA_TSE12} for more details).
%
\EVOSUITE supports multiple different coverage criteria, which can be optimized
at the same time. The default configuration combines branch coverage with
mutation testing~\cite{emse14_mutation} and other basic
criteria~\cite{rojas2015combining}.
%
To cope with the potentially large number of coverage objectives,
DynaMOSA prioritizes them during the search according to their structural
dependencies in the control dependency graph. Initially, the search focuses on
coverage objectives positioned higher in the hierarchy, and the remaining
objectives are incrementally inserted in later generations when their dominator
requirements are covered.

After the search has used up the available search budget (or alternatively, has
achieved 100\% coverage of all coverage objectives), \EVOSUITE applies various
post-search optimizations aimed to improve the readability of the generated
tests~\cite{FrA11c,FrA13a}, such as minimization and addition of test assertions using mutation
analysis~\cite{10.1109/TSE.2011.93}. It also checks all generated tests for compile errors (which may be the result of bugs in \EVOSUITE) or flakiness caused by non-determinism in the class under test, not covered by \EVOSUITE's extensive instrumentation.

\EVOSUITE has been evaluated on open
source as well as industrial software in terms of code
coverage~\cite{fraser2014large,emse_archive,ea_evaluation,dynamosa},
fault-finding
effectiveness~\cite{shamshiri2015automatically,moein2017}, and effects
on developer productivity~\cite{TOSEM_userstudy,ISSTA15_Study} and
software maintenance~\cite{ICST2018_Maintenance}.
%
\EVOSUITE has a longstanding record of success at the unit testing tool
competition, having ranked second in the third edition of the
competition~\cite{evosuiteAtSbst2015} and first in all the other
editions~\cite{evosuiteAtSbst2013,evosuiteAtFittest2013,evosuiteAtSbst2016,evosuiteAtSbst2017,evosuiteAtSbst2018}. In the 2019 SBST contest, \EVOSUITE
achieved the highest overall score of all participating tools~\cite{evosuiteAtSbst2019}, 
although some bugs inhibited its performance.

\section{Tool Setup}

Similar to the previous years, the configuration of \EVOSUITE for the 2020
competition is largely based on its default values since these have been tuned
extensively~\cite{arcuri2013parameter}. 
%
The search algorithm used was DynaMOSA~\cite{dynamosa}, optimizing for the
default set of coverage criteria~\cite{rojas2015combining} (i.e., line
coverage, branch coverage, branch coverage by direct method invocations, weak
mutation testing, output coverage, exception coverage).
%
Other features enabled by default include the use of frequency-based weighted
constants for seeding~\cite{sakti2015instance} as well as support for Java
Enterprise Edition features~\cite{arcuri2016java}. In the case of difficult dependencies and branches that cannot be covered, \EVOSUITE can start using mock objects once a certain percentage of the search
budget has passed~\cite{ICST_Mocking17} .

No major new features have been introduced in \EVOSUITE since the 2019 competition, but several bugs that affected \EVOSUITE's performance during the previous competition have been fixed. In particular, there were several problematic cases when \EVOSUITE was run using very small search budgets.

The test minimization step used by \EVOSUITE is computationally expensive and
sometimes omitted for empirical studies; in the competition, we always enable
the post-processing step of test minimization, because minimized tests are less
likely to break or expose flakiness. However, we aimed to reduce the
post-processing time by including all regression assertions rather than
filtering them with mutation analysis~\cite{10.1109/TSE.2011.93}. While this
makes test cases less readable and potentially more brittle with respect to
future changes in the software under test, neither of these aspects is
evaluated as part of the SBST contest.

\EVOSUITE uses different phases (e.g., initialization, search, minimization,
assertion generation, compilation check, removal of flaky tests). Like in
previous competitions (e.g.,~\cite{evosuiteAtSbst2016}), we allocated 50\% of
the overall time set by the competition organizers for the search, and
distributed the other 50\% equally to the remaining phases.

\section{Benchmark Results}

\begin{table*}[t!]
  \centering
  \caption{\label{table:results}Detailed results of \EVOSUITE on the
    SBST benchmark classes.}
%\vspace{-1em}
\resizebox{0.9\textwidth}{!}{  
\input{mainTable.tex}
}	
\end{table*}

Table~\ref{table:results} summarizes the results achieved by \EVOSUITE on the competition classes and search budgets. In general, the performance of \EVOSUITE was in line with previous
results. As usual, there are some notable cases where \EVOSUITE did not perform well. In the following, we discuss these cases, such that future work can address open problems in automated unit test generation.

%% FESCAR-15 0% coverage 0% mutation score

\subsubsection*{FESCAR-15} According to the results provided by the competition
organizers, it seems that the generated test cases do cover some lines of code
but no branches/conditions.
%Looking at the \texttt{logs.log}, \EVOSUITE also reported
%0\% coverage and we could not find any issues in any %log file. 
For further investigation, we ran \EVOSUITE as a stand-alone tool (i.e., not using the benchmark infrastructure) but it was still not capable of covering any line. Through manual investigation, we notice that the CUT manages objects of the class \texttt{ScheduledExecutorService} and \textit{threads}. Generating tests for CUTs that produce or manage threads is challenging and a well-known open problem in the literature~\cite{FrA13a}. \EVOSUITE does not prevent the CUT from spawning new threads, but it (1) forces the use of a wrapper class that leads to deterministic stack traces and thread names, and (2) any spawned threads are joined and removed after a test execution to ensure a clean state for successive test executions, which may consume a substantial amount of time.

%% FESCAR-2 0% coverage 0% mutation score

\subsubsection*{FESCAR-2} This CUT contains only a few branches (i.e., 14) and lines (i.e., 19). \EVOSUITE was only able to generate test cases that cover around 20\%-25\% of lines but 0\% of condition coverage. The generated tests cover branchless methods and the root branch of the method \texttt{lookup}, which is the only method with some conditions. It is worth noticing that \EVOSUITE could not complete a single generation, even with a search budget of 180s. The CUT is particularly expensive, and each generated test requires seconds for its execution. Low search budgets are thus not feasible nor recommended for this CUT.

\subsubsection*{FESCAR-41, FESCAR-6} \EVOSUITE reached a very low line and branch coverage for these two CUTs, independently from the search budget. These CUT also manage threads, and more precisely objects of the classes \texttt{ThreadPoolExecutor} and \texttt{NamedThreadFactory}.

%% FESCAR-37 86.8% coverage 0% mutation score
%% Annibale: from the logs, I see a \texttt{java.util.concurrent.ExecutionException}.
%% \subsubsection*{FESCAR-37} The classpath is incomplete as there is a
%% \texttt{java.lang.}-\texttt{ClassNotFoundException:  io.netty.channel.Channel}.

%% GUAVA-110 0% coverage 0.6% mutation score

\subsubsection*{GUAVA-110, GUAVA-240} Both classes make use of the annotation \texttt{@Nullable}. GUAVA-110 uses the annotation for the input parameter of the method \texttt{equals(...)}, while GUAVA-240 uses it for the methods \texttt{contains(...)} and \texttt{remove(...)}. The static analyzer implemented in \EVOSUITE does not handle the annotation \texttt{@Nullable} and triggers the error \textit{``Cannot find symbol''}. This, in retrospect, would have been easy to avoid.

\subsubsection*{GUAVA-47} For this class, \EVOSUITE reached very low coverage (both line and branch) and mutation score. To have a better understanding of the underlying issue, we ran \EVOSUITE stand-alone. We notice that our tool could not generate any single test cases, never completing the initialization process (and the generation of the initial population, in particular) even with a search budget of 180s. A more in-depth analysis is needed to discover the root cause of the issue.

\subsubsection*{PDFBOX-157} \EVOSUITE crashed in 9 out ten times with a search budget of 60s, and in 10 out of 10 runs for larger budgets. The exact reasons are not yet clear, and will require further debugging.

%% FESCAR-7 73.3% coverage 0% mutation score
%% GUAVA-212 77.6% coverage 0% mutation score
%% GUAVA-22 51.8% coverage 0% mutation score
%% PDFBOX-265 51.8% coverage 0% mutation score

\subsubsection*{FESCAR-23, FESCAR-37, FESCAR-7, FESCAR-8, GUAVA-212, GUAVA-22,
PDFBOX-265, PDFBOX-278, PDFBOX-285, PDFBOX-40, PDFBOX-91, PDFBOX-62, SPOON-105,
SPOON-253} For all these classes the mutation analysis failed due to a
\texttt{java.util.concurrent.\\ExecutionException} thrown by the experimental
infrastructure, rather than by \EVOSUITE.

%% PDFBOX-235 0% coverage 0% mutation score
%% SPOON-155 0% coverage 0% mutation score

\subsubsection*{PDFBOX-235, SPOON-155} In these two cases, \EVOSUITE crashed while attempting to set up mock objects; these are bugs in \EVOSUITE.

\section{Conclusions}

This paper reports on the participation of the \EVOSUITE test
generation tool in the 8th SBST Java Unit Testing Tool Contest. With
an overall score of \TOTALPOINTS points, \EVOSUITE achieved the
highest score of all tools in the competition. Despite the many years of development, the benchmark used in the competition points out several opportunities for improvement, which we discussed in this paper.

To learn more about \EVOSUITE, visit our Web site:
\begin{center}
%\url{http://evosuite.org/}
\texttt{http://www.evosuite.org}
\end{center}

\textbf{Acknowledgments:} Many thanks to all the contributors to
\EVOSUITE.  This project has been funded by the EPSRC project
``GREATEST'' (EP/N023978/2).

\balance
\bibliographystyle{ACM-Reference-Format}
\bibliography{papers}

\end{document}